---
title: "Create DM table as csv file"
author: "mja@statgroup.dk"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    theme: united
  pdf_document:
    toc: true
    highlight: zenburn
  md_document:
    variant: markdown_github
---

# Introduction

This is a somewhat convoluted script. The idea is to take an existing
cube structure and derive the results. That was usefull initially to
extend a cube. Now, it is not as usefull. So the script may seem
pointless.  TODO(mja): fix it.

# Create DM sample table as CSV file and other files

This script creates the result and codelist for a simple DM table.

```{r}
library(rrdfancillary)
devtools::load_all(pkg="../..")
```

## Get the data and prepare for derivation of summary statistics

```{r}

library(foreign)
library(sqldf)


fnadsl<- system.file("extdata/sample-xpt", "adsl.xpt", package="rrdfqbcrndex")
print(fnadsl)
if (!file.exists(fnadsl)) {
  fnadsl<- file.path("..", "extdata/sample-xpt", "adsl.xpt")
  }
if (!file.exists(fnadsl)) {
  stop("File does not exist - ",fnadsl)
}
adsl<- read.xport(fnadsl)
adsl$TRT01A<- as.character(adsl$TRT01A)
adsl$RACE<- as.character(adsl$RACE)
adsl$SAFFL<- as.character(adsl$SAFFL)
adsl$SEX<- as.character(adsl$SEX)

## SASxport package maps characters and dates etc into more R like data type
## install.packages("SASxport")
## library(SASxport)
## adsl<- as.data.frame(read.xport(fnadsl,as.is=TRUE))
## str(adsl)
```


## Create frame for cube from an existing RDF data cube

The code input a turtle file with an RDF data cube. SQL statements for
calculating the measurements are derived from the cube, and used to
derive the summary statistics. Note: the SQL statements does not show
records where the combination of values lead to 0 observations. This
is handled below, in a not so clever way. A better approach would be
to include the concept of a skeleton in the SQL statements.

ToDo(MJA): move this to rrdfqbcrndcheck or move to another package, like rrdfqcbcrnd0

```{r,eval=FALSE}

library(rrdfqbcrndcheck)

dataCubeFile<- system.file("extdata/sample-rdf", "DC-DM-sample.ttl", package="rrdfqbcrndex")
checkCube <- new.rdf(ontology=FALSE)  # Initialize
load.rdf(dataCubeFile, format="TURTLE", appendTo= checkCube)
summarize.rdf(checkCube)

stmtSQL<- GetSQLFromCube(checkCube) 

cat(stmtSQL$summStatSQL) 

adsl.summ.stat.res<- sqldf( stmtSQL$summStatSQL)
names(adsl.summ.stat.res)<- tolower(gsub("(a|b)\\.","", names(adsl.summ.stat.res)))

```

## Store the SQL statements to a file

```{r, eval=FALSE}

res.text<- stmtSQL$summStatSQL

cr.text<-   paste0("create table qbframe ", "(", paste(names(stmtSQL$qbframe), "TEXT", collapse=", "), ")",";")

in.text<-   paste0(
  paste(
  paste0("insert into qbframe ", "(", paste0(names(stmtSQL$qbframe),collapse=","), ")\n" ),
  "values\n",
  paste0( "(", apply(stmtSQL$qbframe,1,function(x) {paste0('"',x,'"', collapse=",")}), ")",  collapse=",\n"),
  collapse="\n"
  ),";\n")

se.text<- "select * from qbframe;"

tempfile<- file.path(tempdir(),"temp-code.R")
cat(paste('res.text<- "', res.text,'"\n',collapse="\n"), file=tempfile)
cat(paste("cr.text<- '", cr.text,"'\n",collapse="\n"), file=tempfile,append=TRUE)
cat(paste("in.text<- '", in.text,"'\n",collapse="\n"), file=tempfile,append=TRUE)
cat(paste("se.text<- '", se.text,"'\n",collapse="\n"), file=tempfile,append=TRUE)
print(tempfile)

```

## Define SQL statements directly

The statements below are inserted from the file generated above.

Work-around: add SELECT statments below corresponding to the desired statistics.
Update the .csv file, and re-create the cube. Repeat until done. This is of course not the ideal way; waiting to the formular interface to the cube.

Note: the `GetSQLFromCube` functions in package `rrdfqbcrndcheck` generates the SQL statements from the cube obversvations.

```{r}
res.text<- "
SELECT a.TRT01A, '_ALL_' as RACE, a.SEX, a.SAFFL, 'count' as procedure, 'quantity' as factor, '_ALL_' as denominator, '_NULL_' as unit, count(*) as measure from  adsl  as a   group by  a.TRT01A, a.SEX, a.SAFFL
UNION
SELECT a.TRT01A, '_ALL_' as RACE, '_ALL_' as SEX, a.SAFFL, 'mean' as procedure, 'WEIGHTBL' as factor, '_NULL_' as denominator, 'KG' as unit, avg(WEIGHTBL) as measure from  adsl  as a   group by  a.TRT01A, a.SAFFL
UNION
SELECT a.TRT01A, '_ALL_' as RACE, '_ALL_' as SEX, a.SAFFL, 'max' as procedure, 'WEIGHTBL' as factor, '_NULL_' as denominator, 'KG' as unit, max(WEIGHTBL) as measure from  adsl  as a   group by  a.TRT01A, a.SAFFL
UNION
SELECT a.TRT01A, '_ALL_' as RACE, '_ALL_' as SEX, a.SAFFL, 'max' as procedure, 'AGE' as factor, '_NULL_' as denominator, 'YEARS' as unit, max(AGE) as measure from  adsl  as a   group by  a.TRT01A, a.SAFFL
UNION
SELECT a.TRT01A, '_ALL_' as RACE, '_ALL_' as SEX, a.SAFFL, 'median' as procedure, 'WEIGHTBL' as factor, '_NULL_' as denominator, 'KG' as unit, median(WEIGHTBL) as measure from  adsl  as a   group by  a.TRT01A, a.SAFFL
UNION
SELECT a.TRT01A, a.RACE, '_ALL_' as SEX, a.SAFFL, 'count' as procedure, 'quantity' as factor, '_ALL_' as denominator, '_NULL_' as unit, count(*) as measure from  adsl  as a   group by  a.TRT01A, a.RACE, a.SAFFL
UNION
SELECT a.TRT01A, b.RACE, '_ALL_' as SEX, a.SAFFL, 'percent' as procedure, 'proportion' as factor, 'RACE' as denominator, '_NULL_' as unit, 100*avg(a.RACE=b.RACE) as measure from  adsl  as a , (select distinct RACE from adsl) as b group by  a.TRT01A, b.RACE, a.SAFFL
UNION
SELECT a.TRT01A, '_ALL_' as RACE, '_ALL_' as SEX, a.SAFFL, 'stdev' as procedure, 'AGE' as factor, '_NULL_' as denominator, 'YEARS' as unit, stdev(AGE) as measure from  adsl  as a   group by  a.TRT01A, a.SAFFL
UNION
SELECT a.TRT01A, '_ALL_' as RACE, '_ALL_' as SEX, a.SAFFL, 'min' as procedure, 'AGE' as factor, '_NULL_' as denominator, 'YEARS' as unit, min(AGE) as measure from  adsl  as a   group by  a.TRT01A, a.SAFFL
UNION
SELECT a.TRT01A, '_ALL_' as RACE, b.SEX, a.SAFFL, 'percent' as procedure, 'proportion' as factor, 'SEX' as denominator, '_NULL_' as unit, 100*avg(a.SEX=b.SEX) as measure from  adsl  as a , (select distinct SEX from adsl) as b group by  a.TRT01A, b.SEX, a.SAFFL
UNION
SELECT a.TRT01A, '_ALL_' as RACE, '_ALL_' as SEX, a.SAFFL, 'min' as procedure, 'WEIGHTBL' as factor, '_NULL_' as denominator, 'KG' as unit, min(WEIGHTBL) as measure from  adsl  as a   group by  a.TRT01A, a.SAFFL
UNION
SELECT '_ALL_' as TRT01A, a.RACE, '_ALL_' as SEX, a.SAFFL, 'count' as procedure, 'quantity' as factor, '_ALL_' as denominator, '_NULL_' as unit, count(*) as measure from  adsl  as a   group by  a.RACE, a.SAFFL
UNION
SELECT a.TRT01A, '_ALL_' as RACE, '_ALL_' as SEX, a.SAFFL, 'mean' as procedure, 'AGE' as factor, '_NULL_' as denominator, 'YEARS' as unit, avg(AGE) as measure from  adsl  as a   group by  a.TRT01A, a.SAFFL
UNION
SELECT '_ALL_' as TRT01A, '_ALL_' as RACE, a.SEX, a.SAFFL, 'count' as procedure, 'quantity' as factor, '_ALL_' as denominator, '_NULL_' as unit, count(*) as measure from  adsl  as a   group by  a.SEX, a.SAFFL
UNION
SELECT a.TRT01A, '_ALL_' as RACE, '_ALL_' as SEX, a.SAFFL, 'median' as procedure, 'AGE' as factor, '_NULL_' as denominator, 'YEARS' as unit, median(AGE) as measure from  adsl  as a   group by  a.TRT01A, a.SAFFL
UNION
SELECT a.TRT01A, '_ALL_' as RACE, '_ALL_' as SEX, a.SAFFL, 'count' as procedure, 'quantity' as factor, '_ALL_' as denominator, '_NULL_' as unit, count(*) as measure from  adsl  as a   group by  a.TRT01A, a.SAFFL
UNION
SELECT a.TRT01A, '_ALL_' as RACE, '_ALL_' as SEX, a.SAFFL, 'stdev' as procedure, 'WEIGHTBL' as factor, '_NULL_' as denominator, 'KG' as unit, stdev(WEIGHTBL) as measure from  adsl  as a   group by  a.TRT01A, a.SAFFL
"


cr.text<- '
create table qbframe (trt01a TEXT, race TEXT, factor TEXT, procedure TEXT, sex TEXT, saffl TEXT, unit TEXT, denominator TEXT);
'

in.text<- '
insert into qbframe (trt01a,race,factor,procedure,sex,saffl,unit,denominator)
 values
("Xanomeline High Dose","AMERICAN INDIAN OR ALASKA NATIVE","proportion","percent","_ALL_","Y","_NULL_","RACE"),
("Xanomeline Low Dose","AMERICAN INDIAN OR ALASKA NATIVE","proportion","percent","_ALL_","Y","_NULL_","RACE"),
("Xanomeline Low Dose","BLACK OR AFRICAN AMERICAN","proportion","percent","_ALL_","Y","_NULL_","RACE"),
("Placebo","_ALL_","quantity","count","F","Y","_NULL_","_ALL_"),
("_ALL_","WHITE","quantity","count","_ALL_","Y","_NULL_","_ALL_"),
("Xanomeline Low Dose","_ALL_","AGE","min","_ALL_","Y","YEARS","_NULL_"),
("Xanomeline High Dose","_ALL_","AGE","stdev","_ALL_","Y","YEARS","_NULL_"),
("Xanomeline Low Dose","_ALL_","proportion","percent","M","Y","_NULL_","SEX"),
("Placebo","_ALL_","AGE","stdev","_ALL_","Y","YEARS","_NULL_"),
("Xanomeline High Dose","WHITE","quantity","count","_ALL_","Y","_NULL_","_ALL_"),
("Xanomeline Low Dose","WHITE","quantity","count","_ALL_","Y","_NULL_","_ALL_"),
("Xanomeline High Dose","_ALL_","AGE","max","_ALL_","Y","YEARS","_NULL_"),
("Xanomeline High Dose","AMERICAN INDIAN OR ALASKA NATIVE","quantity","count","_ALL_","Y","_NULL_","_ALL_"),
("Placebo","BLACK OR AFRICAN AMERICAN","quantity","count","_ALL_","Y","_NULL_","_ALL_"),
("Placebo","_ALL_","AGE","max","_ALL_","Y","YEARS","_NULL_"),
("Placebo","AMERICAN INDIAN OR ALASKA NATIVE","proportion","percent","_ALL_","Y","_NULL_","RACE"),
("Placebo","BLACK OR AFRICAN AMERICAN","proportion","percent","_ALL_","Y","_NULL_","RACE"),
("Xanomeline Low Dose","_ALL_","quantity","count","_ALL_","Y","_NULL_","_ALL_"),
("Placebo","_ALL_","quantity","count","M","Y","_NULL_","_ALL_"),
("Placebo","_ALL_","proportion","percent","F","Y","_NULL_","SEX"),
("Xanomeline High Dose","_ALL_","quantity","count","_ALL_","Y","_NULL_","_ALL_"),
("Xanomeline Low Dose","_ALL_","quantity","count","M","Y","_NULL_","_ALL_"),
("Xanomeline High Dose","_ALL_","proportion","percent","F","Y","_NULL_","SEX"),
("Xanomeline Low Dose","_ALL_","quantity","count","F","Y","_NULL_","_ALL_"),
("Placebo","_ALL_","AGE","mean","_ALL_","Y","YEARS","_NULL_"),
("Xanomeline High Dose","_ALL_","AGE","median","_ALL_","Y","YEARS","_NULL_"),
("Placebo","_ALL_","AGE","min","_ALL_","Y","YEARS","_NULL_"),
("_ALL_","BLACK OR AFRICAN AMERICAN","quantity","count","_ALL_","Y","_NULL_","_ALL_"),
("Placebo","_ALL_","AGE","median","_ALL_","Y","YEARS","_NULL_"),
("Xanomeline Low Dose","_ALL_","AGE","stdev","_ALL_","Y","YEARS","_NULL_"),
("Xanomeline Low Dose","AMERICAN INDIAN OR ALASKA NATIVE","quantity","count","_ALL_","Y","_NULL_","_ALL_"),
("Xanomeline Low Dose","BLACK OR AFRICAN AMERICAN","quantity","count","_ALL_","Y","_NULL_","_ALL_"),
("Xanomeline Low Dose","_ALL_","AGE","max","_ALL_","Y","YEARS","_NULL_"),
("Xanomeline High Dose","BLACK OR AFRICAN AMERICAN","proportion","percent","_ALL_","Y","_NULL_","RACE"),
("Xanomeline High Dose","WHITE","proportion","percent","_ALL_","Y","_NULL_","RACE"),
("Xanomeline Low Dose","WHITE","proportion","percent","_ALL_","Y","_NULL_","RACE"),
("_ALL_","_ALL_","quantity","count","M","Y","_NULL_","_ALL_"),
("Xanomeline High Dose","_ALL_","quantity","count","M","Y","_NULL_","_ALL_"),
("_ALL_","_ALL_","quantity","count","F","Y","_NULL_","_ALL_"),
("Xanomeline High Dose","_ALL_","quantity","count","F","Y","_NULL_","_ALL_"),
("Placebo","_ALL_","proportion","percent","M","Y","_NULL_","SEX"),
("Xanomeline Low Dose","_ALL_","proportion","percent","F","Y","_NULL_","SEX"),
("Xanomeline Low Dose","_ALL_","AGE","mean","_ALL_","Y","YEARS","_NULL_"),
("Xanomeline High Dose","_ALL_","proportion","percent","M","Y","_NULL_","SEX"),
("Placebo","AMERICAN INDIAN OR ALASKA NATIVE","quantity","count","_ALL_","Y","_NULL_","_ALL_"),
("Xanomeline High Dose","_ALL_","AGE","mean","_ALL_","Y","YEARS","_NULL_"),
("Xanomeline High Dose","BLACK OR AFRICAN AMERICAN","quantity","count","_ALL_","Y","_NULL_","_ALL_"),
("Placebo","WHITE","quantity","count","_ALL_","Y","_NULL_","_ALL_"),
("Xanomeline Low Dose","_ALL_","AGE","median","_ALL_","Y","YEARS","_NULL_"),
("Xanomeline High Dose","_ALL_","AGE","min","_ALL_","Y","YEARS","_NULL_"),
("_ALL_","AMERICAN INDIAN OR ALASKA NATIVE","quantity","count","_ALL_","Y","_NULL_","_ALL_"),
("Placebo","WHITE","proportion","percent","_ALL_","Y","_NULL_","RACE"),
("Placebo","_ALL_","quantity","count","_ALL_","Y","_NULL_","_ALL_"),
("Xanomeline Low Dose","_ALL_","WEIGHTBL","min","_ALL_","Y","KG","_NULL_"),
("Xanomeline High Dose","_ALL_","WEIGHTBL","stdev","_ALL_","Y","KG","_NULL_"),
("Placebo","_ALL_","WEIGHTBL","stdev","_ALL_","Y","KG","_NULL_"),
("Xanomeline High Dose","_ALL_","WEIGHTBL","max","_ALL_","Y","KG","_NULL_"),
("Placebo","_ALL_","WEIGHTBL","max","_ALL_","Y","KG","_NULL_"),
("Placebo","_ALL_","WEIGHTBL","mean","_ALL_","Y","KG","_NULL_"),
("Xanomeline High Dose","_ALL_","WEIGHTBL","median","_ALL_","Y","KG","_NULL_"),
("Placebo","_ALL_","WEIGHTBL","min","_ALL_","Y","KG","_NULL_"),
("Placebo","_ALL_","WEIGHTBL","median","_ALL_","Y","KG","_NULL_"),
("Xanomeline Low Dose","_ALL_","WEIGHTBL","stdev","_ALL_","Y","KG","_NULL_"),
("Xanomeline Low Dose","_ALL_","WEIGHTBL","max","_ALL_","Y","KG","_NULL_"),
("Xanomeline Low Dose","_ALL_","WEIGHTBL","mean","_ALL_","Y","KG","_NULL_"),
("Xanomeline High Dose","_ALL_","WEIGHTBL","mean","_ALL_","Y","KG","_NULL_"),
("Xanomeline Low Dose","_ALL_","WEIGHTBL","median","_ALL_","Y","KG","_NULL_"),
("Xanomeline High Dose","_ALL_","WEIGHTBL","min","_ALL_","Y","KG","_NULL_")
;
'
se.text<-'
select * from qbframe;
'

```

### Evaluate the SQL code

```{r}

adsl.summ.stat.res<- sqldf( res.text )
# adsl.summ.stat$unit<- "_NULL_"
names(adsl.summ.stat.res)<- tolower(gsub("(a|b)\\.","", names(adsl.summ.stat.res)))

rm(qbframe)
sqldf()
sqldf(cr.text )
sqldf(in.text  )
qbframe<- sqldf(se.text)
sqldf()
# str(qbframe)

```

### Combine generated results with the cube frame and write CSV file

```{r}

adsl.summ.stat<- merge(qbframe,adsl.summ.stat.res,by=names(qbframe),all=TRUE)
# adsl.summ.stat<- merge(stmtSQL$qbframe,adsl.summ.stat.res,all=TRUE)
adsl.summ.stat$measure[ is.na(adsl.summ.stat$measure) & adsl.summ.stat$procedure=="count" ]<- 0
adsl.summ.stat

dmtableFile<- file.path( system.file("extdata/sample-cfg", package="rrdfqbcrndex"), "dm.AR.csv" )
## dmtableFile<- file.path(tempdir(),"temp-dm.AR.csv")

write.csv(adsl.summ.stat, file=dmtableFile, row.names=FALSE)
cat("Written to ", dmtableFile, "\n")


```
