---
title: "SPARQL scripts for RDF data cubes"
author: "mja@statgroup.dk"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    theme: united
  pdf_document:
    toc: true
    highlight: zenburn
  md_document:
    variant: markdown_github
---

# SPARQL scripts for the demographics cube (DC-DEMO-sample.ttl)

## Setup

First load the package.
```{r, eval=TRUE}
library(knitr)
library(rrdfancillary)
devtools::load_all(pkg="../..")
```

## Internals

The display of SPARQL script in markdown is done by first creating a chunk, and then using the chunk with the highlight engine in knitr. The advantage of this approach is that all formatting is handled by external packages. To make the highlight output work in markdown two blanks has to be added at the end of line according to markdown syntax.
```{r helperfunction}
mdwrite<- function( sparqlStatements, refname ) {
# fn<- file.path(tempdir(), paste0( refname, ".rq" ) )
fn<- file.path(system.file("extdata/sample-rdf", package="rrdfqbcrndex"), paste0( refname, ".rq" ) )
cat( paste0("## @knitr ", refname), gsub("\\n", "  \n", sparqlStatements), sep="  \n", file=fn)
cat(  gsub("\\n", "  \n", sparqlStatements), sep="  \n")
knitr::read_chunk( fn, from=c(1))
invisible(fn)
}
```


## SPARQL scripts

SPARQL scripts can be used to access the RDF triple store.
In the package the scripts are made by a function generating the SPARQL
script. The generated SPARQL scripts are shown here for the
demographics cube in DC-DEMO-sample.ttl.

The turtle file and the scrips are stored in
```{r}
system.file("extdata/sample-xpt", package="rrdfqbcrndex")
```

## Setup for generating SPARQL scripts for the demographics cube (DC-DEMO-sample.TTL)

The DEMO data exists as a turtle file in the sample-rdf directory.

```{r}
dataCubeFile<- system.file("extdata/sample-rdf", "DC-DEMO-sample.ttl", package="rrdfqbcrndex")
store <- new.rdf()  # Initialize
cat("Loading ", dataCubeFile, "\n")
temp<-load.rdf(dataCubeFile, format="TURTLE", appendTo= store)
summarize.rdf(store)
```

For the functions in the package the datasets definition in the cube is needed.
```{r}
dsdName<- GetDsdNameFromCube( store )
domainName<- GetDomainNameFromCube( store )
forsparqlprefix<- GetForSparqlPrefix( domainName )
```

## SPARQL query for codelists in RDF data cube

The SPARQL query for the dimensions is made by the function GetDimensionsSparqlQuery.
```{r}
codelistRq <- GetCodeListSparqlQueryFull( forsparqlprefix, dsdName )
mdwrite( codelistRq, "DEMOcodelistfull" )
```

```{r DEMOcodelist, results='asis', engine='highlight', engine.opts='-S n3 --inline-css'}
```    

Executing the SPARQL query gives:

```{r}
codelists<- sparql.rdf(store, codelistRq)
knitr::kable(codelists,caption="Codelists")
```


## SPARQL query for dimensions in RDF data cube

The SPARQL query for the dimensions is made by the function GetDimensionsSparqlQuery.
```{r}
dimensionsRq <- GetDimensionsSparqlQuery( forsparqlprefix )
mdwrite( dimensionsRq, "DEMOdimensions" )
```

```{r DEMOdimensions, results='asis', engine='highlight', engine.opts='-S n3 --inline-css'}
```    

Executing the SPARQL query gives:

```{r}
dimensions<- sparql.rdf(store, dimensionsRq)
knitr::kable(dimensions,col.names=c("dimension"),caption="Dimensions")
```

```{r}
store <- new.rdf(ontology=FALSE)



SPARQLinsert<-'
INSERT {  
   ?dimension qb:codeList ?c .  
   ?c skos:prefLabel ?cprefLabel .   
   ?c skos:hasTopConcept ?cl .  
  ?cl skos:prefLabel ?clprefLabel .

WHERE {  
 values ( ?dimension ?c ?cpreflabel ?cl ?clpreflabel ) {  

( crnd-dimension:race  code:race  "Codelist scheme: race"  code:race-AMERICAN_INDIAN_OR_ALASKA_NATIVE          "AMERICAN INDIAN OR ALASKA NATIVE"       )
( crnd-dimension:race  code:race  "Codelist scheme: race"  code:race-ASIAN                                     "ASIAN"                                  )
( crnd-dimension:race  code:race  "Codelist scheme: race"  code:race-BLACK_OR_AFRICAN_AMERICAN                 "BLACK OR AFRICAN AMERICAN"              )
( crnd-dimension:race  code:race  "Codelist scheme: race"  code:race-NATIVE_HAWAIIAN_OR_OTHER_PACIFIC_ISLANDER "NATIVE HAWAIIAN OR OTHER PACIFIC ISLANDER"  )
( crnd-dimension:race  code:race  "Codelist scheme: race"  code:race-WHITE                                     "WHITE"                                  )
( crnd-dimension:race  code:race  "Codelist scheme: race"  code:race-_ALL_                                     "_ALL_"                                  )
( crnd-dimension:race  code:race  "Codelist scheme: race"  code:race-_NONMISS_                                 "_NONMISS_"                              )

 }   
}  
'

update.rdf( store, SPARQLinsert )

cat(asString.rdf(store),"\n")

```
