---
title: "Overview of prefixes for SPARQL"
author: "mja@statgroup.dk"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    theme: united
  pdf_document:
    toc: true
    highlight: zenburn
  md_document:
    variant: markdown_github
---

# Introduction

This vignette provide an overview of the prefixes defined for each of the packages.

Initially the RRDFQBCRND0 packages were built with a not-so-clear understanding. Further componding to the lack of understanding is that `rrdf` function `add.triple` is an interface to Apache/Jena java function addObjectPropery. The addObjectPropery function does not turn a prefixed name into an IRI (using the terminology from [RDF 1.1 Turtle, section 2.4 IRI](https://www.w3.org/TR/turtle/#sec-iri)). This is addressed in the discussion of [Namespaces](http://jena.apache.org/documentation/rdf/#namespaces), see note at end of section

For more understanding on the use of prefixes - in other words: what I would have liked to understand some years ago - see Richard Cyganiaks post [URIs have a namespace part and a local part, right?](http://richard.cyganiak.de/blog/2016/02/uris-have-a-namespace-part-right/).


## Preliminaries
First load the packages.
```{r, eval=TRUE}
library(rrdfqb)
library(rrdfqbcrnd0)
devtools::load_all(pkg="../..")
```

# Prefixes in `rrdfqb`, `rrdfqbcrnd0`, `rrdfqbcrndex`


```{r, eval=TRUE}
rrdfcdisc:::env[["qbCDISCprefixes"]]
```

```{r, eval=TRUE}
rrdfqb:::env[["rrdfqb"]]
```

```{r, eval=TRUE}
rrdfqbcrnd0:::env[["rrdfqbcrnd0"]]
```

These prefixes are also returned by the function rrdfqbcrnd0/R/Get.default.crnd.prefixes.R
```{r, eval=TRUE}
Get.default.crnd.prefixes()
```

The function `GetForSparqlPrefix( domainName )` returns the prefixes for use with an ARMD RDF data cube with the given domainName. Note the domain names is shown in lower case.

```{r, eval=TRUE}
cat(GetForSparqlPrefix( "THE-DOMAIN-NAME" ),"\n")
```

