---
title: "Package overview"
author: "mja@statgroup.dk"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    theme: united
  pdf_document:
    toc: true
    highlight: zenburn
  md_document:
    variant: markdown_github
---

# Introduction

This vignette shows the structure of the `rrdfqbcrnd0` package.

The graphical presentation is made using the R-package `DiagrammeR`
(https://github.com/rich-iannone/DiagrammeR). The R-package uses
`mermaid` a javascript blibrary (https://github.com/knsv/mermaid),
described as in the credits as

> Many thanks to the d3 (http://d3js.org/) and dagre-d3
> (https://github.com/cpettitt/dagre-d3) projects for providing the
> graphical layout and drawing libraries!  Thanks also to the
> js-sequence-diagram (http://bramp.github.io/js-sequence-diagrams)
> project for usage of the grammar for the sequence diagrams.

Note: the html file must be made using `rmarkdown::render`.

This is under development

mja@statgroup.dk 


## Setup

```{r}
knitr::read_chunk("setup.R")
```
```{r setup}
```
```{r, eval=TRUE}
# install.packages("DiagrammeR")
# install.packages("mvbutils")
library(DiagrammeR)
library(mvbutils)
library(rrdf) # apparantly needed 
## library(rrdfqbcrnd0) # should not go here 

pretty.print.rq<- function(rqstring) {
    ## print sparql query with line numbers - use full for finding errors
    ## when using knitr in Rmd files use highlight instead
rqlines<- unlist(strsplit(rqstring,"\n"))
cat(paste( format(seq(rqlines),format="f",digits=floor(log10(length(rqlines)+1))), rqlines, sep=": ", collapse="\n"),"\n")
}

```

# Get package information

```{r, eval=TRUE}
foodweb(where=("package:rrdfqbcrnd0"), cex=0.8, charlim=60,color.lines=FALSE)

funs<- foodweb(where=("package:rrdfqbcrnd0"), cex=0.8, charlim=60,color.lines=FALSE)

model <- new.rdf(ontology=FALSE)
funmat<- funs$funmat
funcPREFIX<- "http://example.org/rrdfqbcrnd0/"
for (r in seq(nrow(funmat))) {
        add.data.triple(model,
                   subject=paste0(funcPREFIX, rownames(funmat)[r]),
                   predicate=paste0(funcPREFIX, "functionName"),
                   data=rownames(funmat)[r]
                   )
    
    for (c in seq(ncol(funmat))) {
        if (funmat[r,c]==1) {
        add.triple(model,
                   subject=paste0(funcPREFIX, rownames(funmat)[r]),
                   predicate=paste0(funcPREFIX, "calls"),
                   object=paste0(funcPREFIX, colnames(funmat)[c])
                   )
    }
    }
    }        

```

# Make the package overview

First get the labels for the functions.
```{r, eval=TRUE}
prefix.text<- "PREFIX prog: <http://example.org/rrdfqbcrnd0/>"
RP.labels.rq<-  paste0( paste0(prefix.text, collapse="\n"), "
select ?rfunc ?rfunclabel
where {
?rfunc prog:functionName ?rfunclabel
    .
    }
", collapse="\n" )
pretty.print.rq(RP.labels.rq)
RP.labels<- data.frame(sparql.rdf(model, RP.labels.rq),stringsAsFactors=FALSE)
knitr::kable(RP.labels)
```

Now get the links between the functions.
```{r, eval=TRUE}
RP.calls.rq<-  paste0( paste0(prefix.text, collapse="\n"), "
select ?rfunc ?rcaller
where {
?rfunc prog:calls ?rcaller
    .
    }
", collapse="\n" )
pretty.print.rq(RP.calls.rq)
RP.calls<- data.frame(sparql.rdf(model, RP.calls.rq),stringsAsFactors=FALSE)
knitr::kable(RP.calls)
```

Finally, generate the `mermaid` commands for showing the process.
```{r, eval=TRUE}
mermaid.commands<- paste(
           "graph TB; ", 
           paste( gsub("^prog:","", gsub("\\.","", RP.labels$rfunc)),"(", RP.labels$rfunclabel, "); ", sep="", collapse=" \n" ),
    paste( gsub("^prog:","", gsub("\\.","", RP.calls$rfunc)), "-->",
          gsub("^prog:","", gsub("\\.","", RP.calls$rcaller)), "; ", sep="", collapse=" \n" ),
          "\n",
          sep=" \n", collapse=" \n")
cat(mermaid.commands,"\n")
# cat(mermaid.commands,"\n", file="mm.mmd")

```

Then display the graph.
```{r, eval=TRUE}
# mermaid("mm.mmd")
DiagrammeR( mermaid.commands )

    
```


# How to make the HTML file

```{r, eval=FALSE}
rmarkdown::render("package-overview.Rmd")
```
