---
title: "RRDF gotcha"
author: "mja@statgroup.dk"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    theme: united
  pdf_document:
    toc: true
    highlight: zenburn
  md_document:
    variant: markdown_github
---

# Setup

First load the package.
```{r, eval=TRUE}
library(rrdf)
library(rrdfancillary)
```

# Getting information out from a store

```{r, eval=TRUE}
storeex<- new.rdf(ontology=FALSE)
query.rq<- '
PREFIX xsd:  <http://www.w3.org/2001/XMLSchema#>
PREFIX ex:   <http://example.org/>
SELECT ?s (isIRI(?s) as ?isiri_s) ?p (isIRI(?p) as ?isiri_p)  ?o (isIRI(?o) as ?isiri_o)  (lang(?o) as ?olang) (datatype(?o) as ?idatatype)
where {
values (?s ?p ?o) {
(ex:a ex:p 1)
(ex:a ex:p 1.0)
(ex:a ex:p "1.00"^^xsd:float)
(ex:a ex:p "1.000"^^xsd:double)
(ex:a ex:p "NaN"^^xsd:float)
(ex:a ex:p "NaN"^^xsd:double)
(ex:b ex:p "a")
(ex:b ex:p "a"@en)
(ex:b ex:p "a"^^xsd:string)
("a" "b" "c")
}
}
'

query.res<- sparql.rdf( storeex, query.rq )
knitr::kable(query.res)
```

# `sparql.rdf` return result is a character matrix

```{r, eval=TRUE}
storeex2<- new.rdf(ontology=FALSE)
query2.rq<- '
PREFIX xsd:  <http://www.w3.org/2001/XMLSchema#>
PREFIX ex:   <http://example.org/>
SELECT ?s ?p  ?o  (datatype(?o) as ?idatatype)
where {
values (?s ?p ?o) {
(ex:a ex:p 1)
(ex:a ex:p 1.0)
(ex:a ex:p 1.0E0)
(ex:a ex:p "NaN"^^xsd:double )
}
}
'

query2.res<- sparql.rdf( storeex2, query2.rq )
knitr::kable(query2.res)

query2.df<- as.data.frame(query2.res, stringsAsFactors=FALSE)
knitr::kable(query2.df)
str(query2.df)
```

So, if the numeric columns are needed, they must explicitely be made by coversion. Below a new column `on` contains the numeric representation of the `o` column.
```{r, eval=TRUE}
query21.df<- query2.df
query21.df$on <- as.numeric(query21.df$o)
knitr::kable(query21.df)
Map(mode,query21.df)
```


# Same triple repeated: difference between Apache/Jena Store and SPARQL insert using Apache/Jena

When the exactly same triples are inserted - only one triple remains.

```{r, eval=TRUE}
store1<- new.rdf(ontology=FALSE)

sparql.rdf( store1, "select ?s ?p ?o (lang(?o) as ?lang) (datatype(?o) as ?datatype) where {?s ?p ?o  }"  )

SPARQLinsert<- '
PREFIX xsd:  <http://www.w3.org/2001/XMLSchema#>
INSERT DATA
{ 
  <http://example.org/subject1>  <http://example.org/property1> "mytext"^^xsd:string .
  <http://example.org/subject1>  <http://example.org/property1> "mytext"^^xsd:string .
}
'
cat(SPARQLinsert,"\n")
update.rdf( store1, SPARQLinsert )
sparql.rdf( store1, "select ?s ?p ?o (lang(?o) as ?lang) (datatype(?o) as ?datatype) where {?s ?p ?o  }"  )
```

Now, of course, it is always better only to store the values once, when it is intended to store one copy.

However, as I thougth that only one trippel is stored, so I was less carefull in some of the code.

Here is what Apache/Jena does when using the RRDF interface.
```{r, eval=TRUE}
store2<- new.rdf(ontology=FALSE)
add.data.triple(
    store2,
    subject="http://example.org/subject1",
    predicate="http://example.org/property1",
    data="mytext",
    lang="en"
)

add.data.triple(
    store2,
    subject="http://example.org/subject1",
    predicate="http://example.org/property1",
    data="mytext",
    type="string"
)
```

Now query the store:
```{r, eval=TRUE}
sparql.rdf( store2, "select ?s ?p ?o where {?s ?p ?o}"  )
```
The two rows look identical. The next query also show language and datatype associate with the object.
```{r, eval=TRUE}
sparql.rdf( store2, "select ?s ?p ?o (lang(?o) as ?lang) (datatype(?o) as ?datatype) where {?s ?p ?o  }"  )
```
The same tripel appears twice! That learned me that the language and data type are important. They make a difference, so to speak.


Now using the same datatype, `string`, gives two triples again.
```{r, eval=TRUE}
store3<- new.rdf(ontology=FALSE)
add.data.triple(
    store3,
    subject="http://example.org/subject1",
    predicate="http://example.org/property1",
    data="mytext",
    type="string"
)

add.data.triple(
    store3,
    subject="http://example.org/subject1",
    predicate="http://example.org/property1",
    data="mytext",
    type="string"
)

sparql.rdf( store3, "select ?s ?p ?o where {?s ?p ?o}"  )
sparql.rdf( store3, "select ?s ?p ?o (lang(?o) as ?lang) (datatype(?o) as ?datatype) where {?s ?p ?o  }"  )
```


Mixing `INSERT DATA` and RRDF `add.data.triple gives` same result - two triples.

```{r, eval=TRUE}
store4<- new.rdf(ontology=FALSE)

SPARQLinsert<- '
PREFIX xsd:  <http://www.w3.org/2001/XMLSchema#>
INSERT DATA
{ 
  <http://example.org/subject1>  <http://example.org/property1> "mytext"^^xsd:string .
}
'
update.rdf( store4, SPARQLinsert )
sparql.rdf( store4, "select ?s ?p ?o (lang(?o) as ?lang) (datatype(?o) as ?datatype) where {?s ?p ?o  }"  )
```
One triple inserted, one triple in the store. Fine!

Now add one triple - exactly the same as the previos.
```{r, eval=TRUE}
add.data.triple(
    store4,
    subject="http://example.org/subject1",
    predicate="http://example.org/property1",
    data="mytext",
    type="string"
)

sparql.rdf( store4, "select ?s ?p ?o (lang(?o) as ?lang) (datatype(?o) as ?datatype) where {?s ?p ?o  }"  )

```
Two triples in the store.

What if doing two `INSERT DATA`?
```{r, eval=TRUE}
store5<- new.rdf(ontology=FALSE)

sparql.rdf( store5, "select ?s ?p ?o (lang(?o) as ?lang) (datatype(?o) as ?datatype) where {?s ?p ?o  }"  )

SPARQLinsert<- '
PREFIX xsd:  <http://www.w3.org/2001/XMLSchema#>
INSERT DATA
{ 
  <http://example.org/subject1>  <http://example.org/property1> "mytext"^^xsd:string .
}
'
update.rdf( store5, SPARQLinsert )
sparql.rdf( store5, "select ?s ?p ?o (lang(?o) as ?lang) (datatype(?o) as ?datatype) where {?s ?p ?o  }"  )
```
One triple - as expected, only triple was inserted.

```{r, eval=TRUE}
update.rdf( store5, SPARQLinsert )
sparql.rdf( store5, "select ?s ?p ?o (lang(?o) as ?lang) (datatype(?o) as ?datatype) where {?s ?p ?o  }"  )
```
One triple - as expected, as the triple already existed.


Lessons learned:

* Apache/Jena interface R to Java behaves differently than Apache Jena handling of Update Scripts.

* Be carefull when changing code.
