---
title: "Reporting process"
author: "mja@statgroup.dk"
date: "`r Sys.Date()`"
always_allow_html: yes
output:
  html_document:
    toc: true
    theme: united
  md_document:
    variant: markdown_github
---


# Introduction

This note shows how to use `RDF` to describe the reporting process, and present it graphically.

The reporting processes are taken from:

* Brega, John, and Colins, Linda. ‘Beyond OpenCDISC: Using Define.xml Metadata to Ensure End-to-End Submission Integrity.’ Gilead offices, Foster City, CA, 2015. (http://www.pharmasug.org/download/sde/sf2015/PharmaSUG_SF2015SDE_07_Brega_Collins.pdf)
* Williams, Tim. ‘Diagram of the SW Initiatives?’ CS Working Group: End to End Connectivity, 7 April 2015. https://phuse.teamworkpm.net/notebooks/48268.

The graphical presentation is made using the R-package `DiagrammeR` (https://github.com/rich-iannone/DiagrammeR). The R-package uses `mermaid` a javascript blibrary (https://github.com/knsv/mermaid), described as in the credits as
> Many thanks to the d3 (http://d3js.org/) and dagre-d3 (https://github.com/cpettitt/dagre-d3) projects for providing the graphical layout and drawing libraries!
> Thanks also to the js-sequence-diagram (http://bramp.github.io/js-sequence-diagrams) project for usage of the grammar for the sequence diagrams.

It may be nesccesary to install the latest version of R-package `DiagrammeR` from GitHub 
```{r,eval=FALSE}
devtools::install_github('rich-iannone/DiagrammeR')
```

This is under development

mja@statgroup.dk 16-jan-2015.


## Setup

```{r}
library(rrdfancillary)
devtools::load_all(pkg="../..")
```

```{r, eval=TRUE}

library("DiagrammeR")
library(rrdf)

pretty.print.rq<- function(rqstring) {
    ## print sparql query with line numbers - use full for finding errors
    ## when using knitr in Rmd files use highlight instead
rqlines<- unlist(strsplit(rqstring,"\n"))
cat(paste( format(seq(rqlines),format="f",digits=floor(log10(length(rqlines)+1))), rqlines, sep=": ", collapse="\n"),"\n")
}

```


# Get the process description

The process description is converted into RDF. Please comment - this
is my first attempt to model a process in RDF. It is done from
scratch - I am aware of there are more sophiscated ways to do it, and
would appreciate learning more on that approach.


```{r, eval=TRUE}
modelTTLFile<- system.file("extdata/sample-rdf", "clinical-data-process-model.ttl", package="rrdfqbcrndex")
model<- load.rdf(modelTTLFile,"TURTLE")
# identify the prefixes in the turtle file
# The prefixes are part of the RRDF model, but I do not know how to extract it from the model.
ttlfile<- readLines(modelTTLFile)
prefix.text<- gsub("@prefix ([^:]+:) +([^>]+>) *\\. *$", "PREFIX \\1 \\2", ttlfile[grep("^@prefix .+$", ttlfile)])
```

# Reporting process from "Using Define.xml Metadata to Ensure End-to-End Submission Integrity", slide 15

The process steps with labels and shapes are stored in a data frame.
```{r, eval=TRUE}
RP.names.rq<- paste0( paste0(prefix.text, collapse="\n"), "
select ?rstep ?rsteplabel ?rstepshape
where {
    ?rstep a repr:PR .
    optional { ?rstep  rdfs:label ?rsteplabel }
    optional { ?rstep repr:DisplayShape ?rstepshape }
    .
    }
order by ?rstep
", collapse="\n" )

pretty.print.rq(RP.names.rq)
RP.names<- data.frame(sparql.rdf(model, RP.names.rq),stringsAsFactors=FALSE)
knitr::kable(RP.names)
```

For later use, get an overview of the shapes used for reporing.
```{r, eval=TRUE}
RP.shapes.rq<- paste0( paste0(prefix.text, collapse="\n"), "
select distinct ?rstepshape
where {
    ?rstep a repr:PR ;
           repr:DisplayShape ?rstepshape
    .
    }
", collapse="\n" )

pretty.print.rq(RP.shapes.rq)

RP.shapes<- data.frame(sparql.rdf(model, RP.shapes.rq),stringsAsFactors=FALSE)
knitr::kable(RP.shapes)
```

Now get the links between the steps.
```{r, eval=TRUE}
RP.feedsto.rq<-  paste0( paste0(prefix.text, collapse="\n"), "
select ?rstepfrom ?rstepto
where {
?rstepfrom repr:FeedsTo ?rstepto 
    .
    }
", collapse="\n" )
pretty.print.rq(RP.feedsto.rq)
RP.feedsto<- data.frame(sparql.rdf(model, RP.feedsto.rq),stringsAsFactors=FALSE)
knitr::kable(RP.feedsto)
```

Finally, generate the `mermaid` commands for showing the process.
```{r, eval=TRUE}
mermaid.commands<- paste(
           "graph TB", 
           paste( RP.names$rstep,"(", RP.names$rsteplabel, ")", sep="", collapse=" \n" ),
          paste( RP.feedsto$rstepfrom, "-->", RP.feedsto$rstepto, sep="", collapse=" \n" ),
          "\n",
          sep=" \n", collapse=" \n")
cat(mermaid.commands,"\n")
```

Then display the graph.
```{r, eval=TRUE}
DiagrammeR( mermaid.commands )
```

# Grid for "Diagram of the SW Initiatives"

The ordering of the rows and columns in the grid is described in the turtle file by a collection.
The Jena extension `list:index` is used to get the position in the collection.
See (http://stackoverflow.com/questions/17523804/is-it-possible-to-get-the-position-of-an-element-in-an-rdf-collection-in-sparql/17530689#17530689).


```{r, eval=TRUE}
wg.names.rq<- paste0( paste0(prefix.text, collapse="\n"), "
PREFIX list: <http://jena.hpl.hp.com/ARQ/list#>
select ?wg ?wglabel ?wgdisplayorder
where {
    ?wg a cswg:WG ;
    rdfs:label ?wglabel .
    optional{
    cswg:CSWGCollection cswg:DisplayOrder ?lswg .
    ?lswg list:index (?wgdisplayorder ?wg)
    }
    }
order by ?wgdisplayorder
", collapse="\n" )

pretty.print.rq(wg.names.rq)

wg.names<- data.frame(sparql.rdf(model, wg.names.rq),stringsAsFactors=FALSE)

lc.names.rq<- paste0( paste0(prefix.text, collapse="\n"), "
PREFIX list: <http://jena.hpl.hp.com/ARQ/list#>
select ?lc ?lclabel ?lcdisplayorder
where {
    ?lc a cswg:DaLi ;
        rdfs:label ?lclabel .
    optional{
          cswg:LifeCycleCollection cswg:DisplayOrder ?lslc .
          ?lslc list:index (?lcdisplayorder ?lc)
          }
    }
order by ?lcdisplayorder
", collapse="\n" )

pretty.print.rq(lc.names.rq)

lc.names<- data.frame(sparql.rdf(model, lc.names.rq),stringsAsFactors=FALSE)
str(lc.names$lcdisplayorder)

grid.rq<- paste0( paste0(prefix.text, collapse="\n"), "
PREFIX list: <http://jena.hpl.hp.com/ARQ/list#>
select ?wg ?wglabel ?wgdisplayorder ?lc ?lclabel ?lcdisplayorder ?cell
where {
{ select ?wg ?wglabel ?wgdisplayorder where {    ?wg a cswg:WG ;
    rdfs:label ?wglabel .
    optional{
    cswg:CSWGCollection cswg:DisplayOrder ?lswg .
    ?lswg list:index (?wgdisplayorder ?wg)
    }
    } }
{ select ?lc ?lclabel ?lcdisplayorder where {
    ?lc a cswg:DaLi ;
        rdfs:label ?lclabel .
    optional{
          cswg:LifeCycleCollection cswg:DisplayOrder ?lslc .
          ?lslc list:index (?lcdisplayorder ?lc)
          }
    }
}
bind( if( exists { ?wg cswg:RelatesTo ?lc }, '*', ' ' ) AS ?cell )
    }
order by ?wgdisplayorder ?lcdisplayorder 
", collapse="\n" )

pretty.print.rq(grid.rq)
```

Finally, create a grid and fill in the values into the grid

```{r, eval=TRUE}
grid<- data.frame(sparql.rdf(model, grid.rq),stringsAsFactors=FALSE)
wglcgrid<- data.frame(matrix(c("?"), nrow= nrow(wg.names), ncol=nrow(lc.names)),
           stringsAsFactors=FALSE, row.names=wg.names$wglabel)
colnames(wglcgrid)<- lc.names$lclabel
for (i in seq(nrow(grid)) ) {
    wglcgrid[as.numeric(grid[i,"wgdisplayorder"])+1,as.numeric(grid[i,"lcdisplayorder"])+1] <- grid[i, "cell"]
}
knitr::kable(wglcgrid)
```


# How to make the HTML file

```{r, eval=FALSE}
rmarkdown::render("reporting-process.Rmd")
```
