---
title: "Using Python to make SPARQL queries"
author: "mja@statgroup.dk"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    theme: united
  pdf_document:
    toc: true
    highlight: zenburn
  md_document:
    variant: markdown_github
---

# Using Python with SPARQL scripts for the demographics cube (DC-DEMO-sample.ttl)

The examples below uses `rdflib` from (https://github.com/RDFLib/rdflib).

As of 28-jun-2015 the code is still in development.

## Requirement

The python package `rdflib` must be installed, see description at (https://github.com/RDFLib/rdflib).

## Get dimensions
```{r, engine='python'}
from rdflib import Graph
g = Graph()
g.parse("../extdata/sample-rdf/DC-DEMO-sample.ttl", format="turtle")
with open ("../extdata/sample-rdf/DEMOdimensions.rq", "r") as myfile:
    dimensionsQuery=myfile.read()
qres = g.query(dimensionsQuery)
for row in qres:
    print("%s" %row)
```

## Get attributes

```{r, engine='python'}
from rdflib import Graph
g = Graph()
g.parse("../extdata/sample-rdf/DC-DEMO-sample.ttl", format="turtle")
with open ("../extdata/sample-rdf/DEMOattributes.rq", "r") as myfile:
    attributesQuery=myfile.read()
qres = g.query(attributesQuery)
for row in qres:
    print("%s" %row)
```

## Get observations

The SPARQL script shows for each observation the dimension, attributes and measures in a row.
The python code shows each variable on a line - that is not the right way to do it.
```{r, engine='python'}
from rdflib import Graph
g = Graph()
g.parse("../extdata/sample-rdf/DC-DEMO-sample.ttl", format="turtle")
with open ("../extdata/sample-rdf/DEMOobservations.rq", "r") as myfile:
    observationsQuery=myfile.read()

qres = g.query(observationsQuery)

for row in qres:
   for t in row:
     print("%s" %t)
 
```

## Get definition for all descriptive statistics

The SPARQL script shows how the function definition for the descriptive statistics is stored in the cube.
```{r, engine='python'}
from rdflib import Graph
g = Graph()
g.parse("../extdata/sample-rdf/DC-DEMO-sample.ttl", format="turtle")
with open ("../extdata/sample-rdf/DEMOprocedure.rq", "r") as myfile:
    procedureQuery=myfile.read()
qres = g.query(procedureQuery)
for row in qres:
     print("%s| %s| %s" %row)

```

## Get definition for descriptive statistics median

The SPARQL script shows how the function definition for the descriptive statistics is stored in the cube.
```{r, engine='python'}
from rdflib import Graph
g = Graph()
g.parse("../extdata/sample-rdf/DC-DEMO-sample.ttl", format="turtle")
with open ("../extdata/sample-rdf/DEMOprocedure-median.rq", "r") as myfile:
    proceduremedianQuery=myfile.read()
qres = g.query(proceduremedianQuery)
for row in qres:
     print("%s| %s| %s" %row)

```

## Get information for selection of data

The SPARQL script shows how the information for selecting data for derivation of univariate statistics is present in the cube.
```{r, engine='python'}
from rdflib import Graph
g = Graph()
g.parse("../extdata/sample-rdf/DC-DEMO-sample.ttl", format="turtle")
with open ("../extdata/sample-rdf/DEMOobservations-R-selection.rq", "r") as myfile:
    observationsRselection=myfile.read()
qres = g.query(observationsRselection)
for row in qres:
     print("%s| %s| %s| %s" %row)

```


