---
title: "Graphical display of SPARQL queries with arq, rapper and graphviz"
author: "mja@statgroup.dk"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    theme: united
  pdf_document:
    toc: true
    highlight: zenburn
  md_document:
    variant: markdown_github
---
 
# Graphical display of results from SPARQL queries for the demographics cube (DC-DEMO-sample.ttl)


## Get one observation and display graphically

The SPARQL query returns all triples for one observation.

Note: under windows it may be nesessary to issue `.libPaths( c( .libPaths(), "~/../../Documents/R/win-library/3.2") )`

```{r}
## install.packages("igraph")
library(igraph)
library(visNetwork)
library(rrdflibs)
library(rrdf)
library(rrdfancillary)
library(rrdfqb)
library(rrdfqbcrnd0)
devtools::load_all(pkg="../..")
```

```{r, eval=TRUE}


cubeVocabularyFn<- system.file("extdata/cube-vocabulary-rdf","cube.ttl", package="rrdfqb")

dataCubeFile<- system.file("extdata/sample-rdf", "DC-DEMO-sample.ttl", package="rrdfqbcrndex")

store <- new.rdf()  # Initialize

cat("Loading cube specfication from ", cubeVocabularyFn, "\n")
tmp<- load.rdf(cubeVocabularyFn, format="TURTLE", appendTo= store)

cat("Loading cube from ", dataCubeFile, "\n")
tmp<- load.rdf(dataCubeFile, format="TURTLE", appendTo= store)

summarize.rdf(store)

dsdName<- GetDsdNameFromCube( store )
domainName<- GetDomainNameFromCube( store )
cat("dsdName ", dsdName, ", domainName ", domainName, "\n" )
forsparqlprefix<- GetForSparqlPrefix( domainName )

```

```{r, eval=TRUE}

rqfile<- system.file("extdata/sample-rdf", "OneQBobservationWithClasses.rq", package="rrdfqbcrndex")

# begin: Not used
oneQBobservationrq<- paste0(readLines(rqfile),collapse="\n")
oneQBobservationrq
oneobsStore<- construct.rdf(store, oneQBobservationrq)
# End: not used
```

# Showing the RDF data cube observations

Note: This code just work - can be made simpler and easier to read/use.
Clean up - bad practice to reuse variables without at least setting to NULL.

## Showing one RDF data cube observation

```{r, eval=TRUE}

rqnodes<-  paste( forsparqlprefix,
'select distinct ?id (if(isiri(?id),?id,substr(str(?id),1,10)) as ?label) (?id as ?title) ("box" as ?shape)
 ?group
where {
{select distinct (?s as ?id) (?sc as ?group)
 where
 {?s ?p ?o.
 optional{?s a ?sc}
 FILTER (?sc!= rdfs:Resource && ?sc!=qb:Attachable &&
         ?sc!=owl:Class && ?sc!=rdfs:Class)
}
# values (?s) { (ds:obs014) (ds:obs015) }
values (?s) { (ds:obs015) }
}
UNION
{select distinct (?o as ?id) (if(isLITERAL(?o), "literal", ?so) as ?group) where {?s ?p ?o. optional{?o a ?so. ?so a rdfs:Class.}
 FILTER ( (?so = skos:Concept || isLITERAL(?o))
         && ?p != crnd-attribute:unit
         && ?p != crnd-attribute:denominator
         && ?p != crnd-attribute:rowno
         && ?p != crnd-attribute:colno
         && ?p != crnd-attribute:cellpartno
         && ?p != rdfs:label
         && ?p != crnd-attribute:measurefmt
         )
  } 
# values (?s) { (ds:obs014) (ds:obs015) }
values (?s) { (ds:obs015) }
 }
UNION
{select distinct (?so as ?id) (if(isLITERAL(?o), "literal", ?so) as ?group) where {?s qb:dataSet ?so.
# values (?s) { (ds:obs014) (ds:obs015) }
values (?s) { (ds:obs015) }
 }
}
}
order by ?group
',
"\n"                               
)
nodes<- data.frame(sparql.rdf( store, rqnodes ))
knitr::kable(nodes)


rqedges<-  paste( forsparqlprefix,
'select distinct (?s as ?from) (?o as ?to) (?p as ?label)
where
{ ?s ?p ?o
# values (?s) { (ds:obs014) (ds:obs015) }
values (?s) { (ds:obs015) }
}
',
"\n"                               
)
edges<- data.frame(sparql.rdf( store, rqedges ))
knitr::kable(edges)

visNetwork(nodes, edges, height = "1000px", width = "100%") 
# visNetwork(nodes, edges, height = "1000px", width = "100%") %>% visIgraphLayout(layout = "layout_as_tree")

   
``` 


## Showing two RDF data cube observation

```{r, eval=TRUE}

rqnodes<-  paste( forsparqlprefix,
'select distinct ?id (if(isiri(?id),?id,substr(str(?id),1,10)) as ?label) (?id as ?title) ("box" as ?shape)
 ?group
where {
{select distinct (?s as ?id) (?sc as ?group)
 where
 {?s ?p ?o.
 optional{?s a ?sc}
 FILTER (?sc!= rdfs:Resource && ?sc!=qb:Attachable &&
         ?sc!=owl:Class && ?sc!=rdfs:Class)
}
# values (?s) { (ds:obs014) (ds:obs015) }
values (?s) { (ds:obs015) }
}
UNION
{select distinct (?o as ?id) (if(isLITERAL(?o), "literal", ?so) as ?group) where {?s ?p ?o. optional{?o a ?so. ?so a rdfs:Class.}
 FILTER ( (?so = skos:Concept || isLITERAL(?o))
         && ?p != crnd-attribute:unit
         && ?p != crnd-attribute:denominator
         && ?p != crnd-attribute:rowno
         && ?p != crnd-attribute:colno
         && ?p != crnd-attribute:cellpartno
         && ?p != rdfs:label
         && ?p != crnd-attribute:measurefmt
         )
  } 
values (?s) { (ds:obs014) (ds:obs015) }
 }
UNION
{select distinct (?so as ?id) (if(isLITERAL(?o), "literal", ?so) as ?group) where {?s qb:dataSet ?so.
values (?s) { (ds:obs014) (ds:obs015) }
 }
}
}
order by ?group
',
"\n"                               
)
nodes<- data.frame(sparql.rdf( store, rqnodes ))
knitr::kable(nodes)


rqedges<-  paste( forsparqlprefix,
'select distinct (?s as ?from) (?o as ?to) (?p as ?label)
where
{ ?s ?p ?o
values (?s) { (ds:obs014) (ds:obs015) }
}
',
"\n"                               
)
edges<- data.frame(sparql.rdf( store, rqedges ))
knitr::kable(edges)

visNetwork(nodes, edges, height = "1000px", width = "100%") 
# visNetwork(nodes, edges, height = "1000px", width = "100%") %>% visIgraphLayout(layout = "layout_as_tree")

   
``` 

## Showing all RDF data cube observation

```{r, eval=TRUE}

rqnodes<-  paste( forsparqlprefix, '
select distinct ?id (if(isiri(?id),?id,substr(str(?id),1,10)) as ?label) (?id as ?title) ("box" as ?shape)
 ?group
where {
{select distinct (?s as ?id) (?sc as ?group)
 where
 {?s ?p ?o.
?s a qb:Observation.
 optional{?s a ?sc}
 FILTER (?sc!= rdfs:Resource && ?sc!=qb:Attachable &&
         ?sc!=owl:Class && ?sc!=rdfs:Class)
 }
}
UNION
{select distinct (?o as ?id) (if(isLITERAL(?o), "literal", ?so) as ?group) where {?s ?p ?o. 
?s a qb:Observation.

optional{?o a ?so. ?so a rdfs:Class.}
 FILTER ( (?so = skos:Concept || isLITERAL(?o))
         && ?p != crnd-attribute:unit
         && ?p != crnd-attribute:denominator
         && ?p != crnd-attribute:rowno
         && ?p != crnd-attribute:colno
         && ?p != crnd-attribute:cellpartno
         && ?p != rdfs:label
         && ?p != crnd-attribute:measurefmt
         )
  } 
 }
UNION
{select distinct (?so as ?id) (if(isLITERAL(?o), "literal", ?so) as ?group) where {?s qb:dataSet ?so.
 ?s a qb:Observation.
 }
}
}
order by ?group
',
"\n"                               
)
nodes<- NULL
nodes<- data.frame(sparql.rdf( store, rqnodes ))
knitr::kable(nodes)


rqedges<-  paste( forsparqlprefix,
'select distinct (?s as ?from) (?o as ?to) (?p as ?label)
where
{ ?s ?p ?o .
 ?s a qb:Observation .
}
',
"\n"                               
)
edges<- NULL
edges<- data.frame(sparql.rdf( store, rqedges ))
knitr::kable(edges)

visNetwork(nodes, edges, height = "1000px", width = "100%") 
# visNetwork(nodes, edges, height = "1000px", width = "100%") %>% visIgraphLayout(layout = "layout_as_tree")

   
``` 
