---
title: "Graphical display of SPARQL queries with arq, rapper and graphviz"
author: "mja@statgroup.dk"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    theme: united
  pdf_document:
    toc: true
    highlight: zenburn
  md_document:
    variant: markdown_github
---
 
# Graphical display of results from SPARQL scripts for the demographics cube (DC-DEMO-sample.ttl)

The examples below uses `arq` from Apache Jena
(http://jena.apache.org). To install arq - download and unpack the
latest version of apache-jena from
(http://jena.apache.org/download/index.cgi). Put the executable `arq`
in the path, or invoke `arq` with the full path to the directory with
arq.

The use of arq is described many places, see for example
(http://www.learningsparql.com/).

All `arq` commands below are to be run in the directory with the
sample files, which is `inst/extdata/sample-rdf` directory or
`extdata/sample-rdf` depending on the whether the development version
or the installed version of the package is used.

The cd below in each code block is included because I could not find a
quick way to get the code chunk executed in that directory. knitr is
flexible enough to do it, I have not yet found the right way to do it.
So, ignore the repeated cd ..


## Get one observation and display graphically

The SPARQL query returns all triples for one observation, and stores
it in the file `fordot.ttl`. The file is used as input to `rapper` and
converted to the dot format, and displayed using the `dot` program
part of Graphviz (http://www.graphviz.org/).

```{r, engine='bash'}
cd ../extdata/sample-rdf
arq --data DC-DEMO-sample.ttl --query OneQBobservation.rq > fordot.ttl
cat fordot.ttl
rapper -i turtle -o dot fordot.ttl > fordot.dot
dot -x -Tpdf -ograph.pdf fordot.dot
```

The pdf file can then be viewed using a pdf viewer. 

Alternatively, knitr supports presenting dot as show below.

```{r dot-ex, engine ='dot', fig.cap ='SPARQL query', cache=TRUE}
digraph {
	rankdir = LR;
	charset="utf-8";

	"Rds:obs014" -> "Rqb:Observation" [ label="rdf:type" ];
	"Rds:obs014" -> "L14" [ label="rdfs:label" ];
	"Rds:obs014" -> "Rds:dataset-DEMO" [ label="qb:dataSet" ];
	"Rds:obs014" -> "L1|Datatype: xsd:string" [ label="crnd-attribute:cellpartno" ];
	"Rds:obs014" -> "L2|Datatype: xsd:string" [ label="crnd-attribute:colno" ];
	"Rds:obs014" -> "L" [ label="crnd-attribute:denominator" ];
	"Rds:obs014" -> "L3|Datatype: xsd:string" [ label="crnd-attribute:rowno" ];
	"Rds:obs014" -> "LNA|Datatype: xsd:string" [ label="crnd-attribute:unit" ];
	"Rds:obs014" -> "Rcode:agegr1-_ALL_" [ label="crnd-dimension:agegr1" ];
	"Rds:obs014" -> "Rcode:ethnic-_ALL_" [ label="crnd-dimension:ethnic" ];
	"Rds:obs014" -> "Rcode:factor-quantity" [ label="crnd-dimension:factor" ];
	"Rds:obs014" -> "Rcode:procedure-count" [ label="crnd-dimension:procedure" ];
	"Rds:obs014" -> "Rcode:race-_ALL_" [ label="crnd-dimension:race" ];
	"Rds:obs014" -> "Rcode:sex-M" [ label="crnd-dimension:sex" ];
	"Rds:obs014" -> "Rcode:trt01a-Xanomeline_Low_Dose" [ label="crnd-dimension:trt01a" ];
	"Rds:obs014" -> "L34|Datatype: xsd:int" [ label="crnd-measure:measure" ];

	// Resources
	"Rds:obs014" [ label="ds:obs014", shape = ellipse, color = blue ];
	"Rqb:Observation" [ label="qb:Observation", shape = ellipse, color = blue ];
	"Rds:dataset-DEMO" [ label="ds:dataset-DEMO", shape = ellipse, color = blue ];
	"Rcode:agegr1-_ALL_" [ label="code:agegr1-_ALL_", shape = ellipse, color = blue ];
	"Rcode:ethnic-_ALL_" [ label="code:ethnic-_ALL_", shape = ellipse, color = blue ];
	"Rcode:factor-quantity" [ label="code:factor-quantity", shape = ellipse, color = blue ];
	"Rcode:procedure-count" [ label="code:procedure-count", shape = ellipse, color = blue ];
	"Rcode:race-_ALL_" [ label="code:race-_ALL_", shape = ellipse, color = blue ];
	"Rcode:sex-M" [ label="code:sex-M", shape = ellipse, color = blue ];
	"Rcode:trt01a-Xanomeline_Low_Dose" [ label="code:trt01a-Xanomeline_Low_Dose", shape = ellipse, color = blue ];

	// Anonymous nodes

	// Literals
	"L14" [ label="14", shape = record ];
	"L1|Datatype: xsd:string" [ label="1|Datatype: xsd:string", shape = record ];
	"L2|Datatype: xsd:string" [ label="2|Datatype: xsd:string", shape = record ];
	"L" [ label="", shape = record ];
	"L3|Datatype: xsd:string" [ label="3|Datatype: xsd:string", shape = record ];
	"LNA|Datatype: xsd:string" [ label="NA|Datatype: xsd:string", shape = record ];
	"L34|Datatype: xsd:int" [ label="34|Datatype: xsd:int", shape = record ];

	label="\n\nModel:\n(Unknown)\n\nNamespaces:\nprov: http://www.w3.org/ns/prov#\ncrnd-dimension: http://www.example.org/dc/dimension#\nmms: http://rdf.cdisc.org/mms#\ncrnd-measure: http://www.example.org/dc/measure#\ncode: http://www.example.org/dc/code/\nqb: http://purl.org/linked-data/cube#\ndccs: http://www.example.org/dc/demo/dccs/\nrdfs: http://www.w3.org/2000/01/rdf-schema#\ndcat: http://www.w3.org/ns/dcat#\npav: http://purl.org/pav\ndct: http://purl.org/dc/terms/\nxsd: http://www.w3.org/2001/XMLSchema#\nowl: http://www.w3.org/2002/07/owl#\nrdf: http://www.w3.org/1999/02/22-rdf-syntax-ns#\nskos: http://www.w3.org/2004/02/skos/core#\ncts: http://rdf.cdisc.org/ct/schema#\nrrdfqbcrnd0: http://www.example.org/rrdfqbcrnd0/\nds: http://www.example.org/dc/demo/ds/\ncrnd-attribute: http://www.example.org/dc/attribute#\n";
}
```

<!-- ```{r cache=FALSE} -->
<!-- knitr::read_chunk("../extdata/sample-rdf/fordot.dot") -->
<!-- ``` -->

<!-- ```{r unnamed-chunk-2, engine ='dot', fig.cap ='SPARQL query'} -->
<!-- ``` -->

<!-- cd ../extdata/sample-rdf -->
<!-- arq --data DC-DEMO-sample.ttl --query OneQBobservation.rq > fordot.ttl -->
<!-- rapper -q -i turtle -o dot fordot.ttl -->
