---
title: "Using NoSPA-RDF-Data-Cube-Validator"
author: "mja@statgroup.dk"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    theme: united
  pdf_document:
    toc: true
    highlight: zenburn
  md_document:
    variant: markdown_github
---

# Validating RDF data cube use NoSPA-RDF-Data-Cube-Validator

The NoSPA-RDF-Data-Cube-Validator is a standalone java implementation
avialable at
(https://github.com/yyz1989/NoSPA-RDF-Data-Cube-Validator). The jar
file is available from
(https://github.com/yyz1989/NoSPA-RDF-Data-Cube-Validator/releases).

The code below assumes the jar file is available at `/opt/NoSPA-RDF-Data-Cube-Validator-jar`.

The cd below in each code block is included because I could not find a
quick way to get the code chunk executed in that directory. knitr is
flexible enough to do it, I have not yet found the right way to do it.
So, ignore the repeated cd ..

## Integrity contstraints check on DC-DM-sample.ttl
```{r, engine='bash'}
cd ../extdata/sample-rdf
java -jar /opt/NoSPA-RDF-Data-Cube-Validator-jar/nospa-rdf-data-cube-validator-0.9.9-jar-with-dependencies.jar DC-DM-sample.ttl nospa
```

## Integrity contstraints check on DC-AE-sample.ttl
```{r, engine='bash'}
cd ../extdata/sample-rdf
java -jar /opt/NoSPA-RDF-Data-Cube-Validator-jar/nospa-rdf-data-cube-validator-0.9.9-jar-with-dependencies.jar DC-AE-sample.ttl nospa
```

## Integrity contstraints check on DC-DEMO-sample.ttl
```{r, engine='bash'}
cd ../extdata/sample-rdf
java -jar /opt/NoSPA-RDF-Data-Cube-Validator-jar/nospa-rdf-data-cube-validator-0.9.9-jar-with-dependencies.jar DC-DEMO-sample.ttl nospa
```

## Display of the validation files
```{r, engine='bash'}
cd ../extdata/sample-rdf
cat validation_result_*.md
```
(This is not a very nice way to do it).
