---
title: "Show the procedure codelist"
author: "PhuseSubTeamAnalysisResults@example.org"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    theme: united
  pdf_document:
    toc: true
    highlight: zenburn
  md_document:
    variant: markdown_github
---

## Introduction
This vignette shows
- the procedure codelist

# Setup

```{r}
library(rrdfancillary)
devtools::load_all(pkg="../..")
library(rrdfqb)
```

# Create RDF data cube
The RDF data cube will be created from two data.frames containing data and metadata.

## Define data
The data are defined as data frame, and the data frame is displayed.
```{r}
thisGetDescrStatProcedure<- GetDescrStatProcedure()
procedureName<- gsub( "code:procedure-", "", names(thisGetDescrStatProcedure))
obsData<- data.frame(
  category=rep( "AA-group", length(procedureName)),
  procedure=procedureName,
  factor=rep( "VARA", length(procedureName)),
  unit=rep( "Not given", length(procedureName)),
  denominator=rep( " ", length(procedureName)),
  measure=seq(length(procedureName)),
  stringsAsFactors=FALSE  )
knitr::kable(obsData)
```

## Define meta data
The metadata used for generating the RDF data cube are also defined as data frame and displayed.
```{r}
cubeMetadata<- data.frame(
  compType=c("dimension", "dimension", "dimension", "unit", "denominator", "measure", "metadata"),
  compName=c("category", "procedure", "factor", "attribute", "attribute", "measure", "domainName"),
  codeType=c("DATA", "DATA", "DATA", " ", " ", "<NA>","<NA>"),
  nciDomainValue=c(" "," "," "," ", " ", " "," "),
  compLabel=c("Category", "Statistical procedure", "Type of procedure", "Result", "Unit", "Denominator", "EXAMPLE"),
  Comment=c(" "," "," "," "," "," "," "),
  stringsAsFactors=FALSE  )
knitr::kable(cubeMetadata)
```

## Create RDF data cube

The RDF data cube for the data above is created using
```{r}
outcube<- BuildCubeFromDataFrames(cubeMetadata, obsData )
```
This shows a simple use of the BuildCubeFromDataFrames function. 
The warning message from log4j can be ignored.

The RDF data cube is serialized in turtle format and stored as a text file in
```{r}
cat(normalizePath(outcube),"\n")
```

# Query the cube using SPARQL

Now take a look at the generated cubes by loading the turle file.

```{r}
dataCubeFile<- outcube
```

The rest of the code only depends on the value of dataCubeFile.
The code demonstrates the use of the rrdf library.

```{r}
cube <- new.rdf()  # Initialize
temp<-load.rdf(dataCubeFile, format="TURTLE", appendTo= cube)
summarize.rdf(cube)
```

The next statements are needed for the current implementation of the cube, and may change in future versions.
```{r}
## TODO: reconsider the use of domain specific prefixes
dsdName<- GetDsdNameFromCube( cube )
domainName<- GetDomainNameFromCube( cube )
cat("dsdName ", dsdName, ", domainName ", domainName, "\n" )

prefixlist<- Get.prefixlist.from.df(GetForSparqlPrefix.as.df(domainName=domainName))
forsparqlprefix<- GetForSparqlPrefix( domainName )
cat(forsparqlprefix,"\n")
```

The variable forsparqlprefix contains the prefix statements applicable
for the present data cube. The use of prefixes makes the SPARQL query
shorter, and more readable. The present version of the package defines
namespaces dccs: and ds: where the domainname is included.
TODO: Consider other approach for including the domainname, or use other concept.

## Add `rdfs:seeAlso` for descriptive statistics

```{r}
        add.triple(cube,
                   paste0(prefixlist$prefixCODE, "procedure-median"),
                   paste0(prefixlist$prefixRDFS,"seeAlso" ),
                   "purl.obolibrary.org/obo/OBI_0200119")
        add.triple(cube,
                   paste0(prefixlist$prefixCODE, "procedure-stddev"),
                   paste0(prefixlist$prefixRDFS,"seeAlso" ),
                   "purl.obolibrary.org/obo/STATO_0000037")
        add.triple(cube,
                   paste0(prefixlist$prefixCODE, "procedure-mean"),
                   paste0(prefixlist$prefixRDFS,"seeAlso" ),
                   "purl.obolibrary.org/obo/IAO_0000125")
```


## Getting all codelists as data.frame

The SPARQL query for codelists are shown in the next output.
```{r, echo=FALSE}
codelists.rq<- GetCodeListSparqlQuery( forsparqlprefix, dsdName )
cat(codelists.rq)
```

Executing the SPARQL query gives the code list as a data frame.
```{r, echo=FALSE}
cube.codelists<- as.data.frame(sparql.rdf(cube, codelists.rq), stringsAsFactors=FALSE)

## TODO instead of gsub make a more straightforward way
## TODO this involves a new version of the ph.recode function
## TODO(mja): change the GetCodeListSparqlQuery to plain text versions of the clc also
cube.codelists$clc<- gsub("code:","",cube.codelists$cl)
knitr::kable(print(cube.codelists[,c("vn", "clc", "clprefLabel")]))

```

## Getting the procedure codelist as data.frame

```{r, echo=FALSE}
procedure.codelists.rq<- paste( forsparqlprefix,
'
select ?s ?p ?o
where { 
?s ?p ?o .
?s a code:Procedure
 }    
',
"\n"
)

cat(procedure.codelists.rq)
```

Executing the SPARQL query gives the code list as a data frame.
```{r, echo=FALSE}
cube.procedure.codelists<- as.data.frame(sparql.rdf(cube, procedure.codelists.rq), stringsAsFactors=FALSE)
knitr::kable(print(cube.procedure.codelists))

```

## Get the procedure codelist as turtle with Stato References

```{r, echo=FALSE}
construct.codelists.rq<- paste( forsparqlprefix,
'
construct { ?s ?p ?o } 
where { 
{ ?s ?p ?o . ?s a code:Procedure. }
union
{ ?s ?p ?o . values(?s) {(code:Procedure)} }
 } 
',
"\n"
)

cat(construct.codelists.rq)
```

```{r, echo=FALSE}
procedure.codelist.store<- construct.rdf(cube, construct.codelists.rq)
targetDir<- system.file("extdata/sample-rdf", package="rrdfqbcrnd0")
procedure.codelist.fn <- file.path(targetDir,"procedure-code-list.ttl" ) # Full path to file
outcodelist<- save.rdf(procedure.codelist.store, procedure.codelist.fn, format="TURTLE")
```

The RDF data cube is serialized in turtle format and stored as a text file in
```{r}
cat(normalizePath(outcodelist),"\n")
```

The triples are:
```{r}
cat(readLines(procedure.codelist.fn),"\n")
```

