---
title: "Store CDISC RDF as RRDF data model"
author: "PhuseSubTeamAnalysisResults@example.org"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    theme: united
  pdf_document:
    toc: true
    highlight: zenburn
  md_document:
    variant: markdown_github

---

# Store CDISC RDF as RRDF data model

##  Preliminaries

The path to the package root is `../..` relative to the present working directory.
```{r}
library(rrdfancillary)
devtools::load_all(pkg="../..")
```

## Directory for storing rdf.cdisc.org and RDF data cube specification

```{r get-rdf-cdisc-org-setup-export.sh, eval=FALSE, engine='bash'}
echo change below to suit your setup
export CDISCRDF_HOME=~/projects/phrmwg
```

## Make local copy of the rdf.cdisc.org
Here is how to make a local copy of the rdf.cdisc.org - with Bash Shell on Linux.
In my setup the `~/projects/phrmwg` directory contains a git pull of [rdf.cdisc.org].

```{r get-rdf-cdisc-org-setup.sh, eval=FALSE, engine='bash'}
echo only if directory does not exist 
mkdir -p ${CDISCRDF_HOME}  
cd ${CDISCRDF_HOME}

git clone https://github.com/phuse-org/rdf.cdisc.org.git

```

## Finding the files

Replace directory for find with the appropriate directory for your setup.
Also change below for the function `Get.filenames.for.cdisc.standards.R`.

This is one way to find the file names:
```
find ${CDISCRDF_HOME} -name "*.rdf" -o -name "*.owl" -o -name "*.ttl"
```

The result of the function call Get.filenames.for.cdisc.standards() in
the rrdfqbcrdn0 package is intended to match the output from find.
```{r, results='asis', eval=FALSE} 
Get.filenames.for.cdisc.standards()
```

Note: this is not a straightforward way to do it.

# R-code

## Setup 
```{r, results='asis', eval=TRUE}
library(rrdf)
library(tools)
```

## Create local version using RRDF from local directory conting rdf.cdisc.org

To conserve space in the R package the files are located outside the package.

The files are stored beneath the directory given by fileStemDirectory.
The path relative to fileStemDirectory is stored in filesToLoad.

```{r, results='asis', eval=TRUE}
fileStemDirectory<- "~/projects/phrmwg"

filesToLoad<- c(
"rdf.cdisc.org/resources/w3.org/skos.rdf", 
"rdf.cdisc.org/resources/dublincore.org/dcam.rdf", 
"rdf.cdisc.org/resources/dublincore.org/dcelements.rdf", 
"rdf.cdisc.org/resources/dublincore.org/dcterms.rdf", 
"rdf.cdisc.org/terminology-2013-06-28/glossary-terminology.owl", 
"rdf.cdisc.org/terminology-2013-06-28/cdash-terminology.owl", 
"rdf.cdisc.org/terminology-2013-06-28/sdtm-terminology.owl", 
"rdf.cdisc.org/terminology-2013-06-28/qs-terminology.owl", 
"rdf.cdisc.org/terminology-2013-06-28/send-terminology.owl", 
"rdf.cdisc.org/terminology-2013-06-28/adam-terminology.owl", 
"rdf.cdisc.org/std/sdtm-1-2.ttl",
"rdf.cdisc.org/std/all-standards.ttl",
"rdf.cdisc.org/std/cdash-1-1.ttl",
"rdf.cdisc.org/std/sdtmig-3-1-3.ttl",
"rdf.cdisc.org/std/sdtmig-3-1-2.ttl",
"rdf.cdisc.org/std/adamig-1-0.ttl",
"rdf.cdisc.org/std/sendig-3-0.ttl",
"rdf.cdisc.org/std/sdtm-1-3.ttl",
"rdf.cdisc.org/std/adam-2-1.ttl", 
"rdf.cdisc.org/schemas/ct-schema.owl", 
"rdf.cdisc.org/schemas/meta-model-schema.owl", 
"rdf.cdisc.org/schemas/cdisc-schema.owl"
  )
```

Then the files are input and the size of the resulting file is reported.

```{r, results='asis', eval=TRUE}
cdisc.save.zip<- Create.cdisc.standards.from.local(
  cdisc.files.dir= fileStemDirectory,
  CDISCfilelist= filesToLoad
  )

cdisc.save.zip.info<- file.info(cdisc.save.zip)
message("File ", normalizePath(cdisc.save.zip), " created ", cdisc.save.zip.info$ctime, " size ", cdisc.save.zip.info$size, " bytes")
```

### Creating SPARQL construct script for getting CDISC standard using FROM dataset

The following is experimental and not needed for the present
package. See below for explanation.

```{r, results='asis', eval=TRUE}

rdf.cdisc.org.URLstem<- "https://github.com/phuse-org/rdf.cdisc.org/raw/master"
rdf.data.cube.URLstem<- "http://publishing-statistical-data.googlecode.com/svn/trunk/specs/src/main/vocab"
rqfilesToLoad<- filesToLoad
rqfilesToLoad<- gsub("rdf.cdisc.org", rdf.cdisc.org.URLstem, rqfilesToLoad )
rqfilesToLoad<- gsub("rdf-data-cube", rdf.data.cube.URLstem, rqfilesToLoad )


SPARQLscript<- paste(
  "CONSTRUCT { ?s ?p ?o }",
paste0( "FROM ", "<", rqfilesToLoad, ">", collapse="\n" )
  ,
  "WHERE { ?s ?p ?o }", sep="\n", collapse="\n"
)

cat(SPARQLscript,"\n")
# change to be part of package
SPARQLscriptfn<- file.path(tempdir(),"get-rdf-disc-org.rq")
writeLines( SPARQLscript, con=SPARQLscriptfn )
cat("SPARQL script stored in  ", normalizePath(SPARQLscriptfn), "\n")

targetDir<- system.file("extdata/CDISC-standards-rdf", package="rrdfcdisc")
targetFile<- file.path(targetDir,basename(SPARQLscriptfn) )

if (file.copy( SPARQLscriptfn, targetFile, overwrite=TRUE)) {
  cat("SPARQL script copied to ", normalizePath(targetFile), "\n")
}


```

ToDo(mja): The same approach could be used for an SPARQL update load script - that could make a load for each file.

### Using the script with R
The R-code below does not work with rrdf.

```{r, results='asis', eval=FALSE}
cdisc.rdf.store <- new.rdf()
results <- construct.rdf(cdisc.rdf.store, SPARQLscript )
summarize.rdf(results)
```

The code below assumes a locate fuseki instance is running.
```{r, results='asis', eval=FALSE}
endpoint<- "http://localhost:3030/arm/query"
results.fuseki <- construct.remote(endpoint, SPARQLscript )
summarize.rdf(results.fuseki)
```
This worked partially with fuseki 2.3 as of 28-oct-2015: 42199 triples was retrieved. The fuseki endpoint was started using the configuration file `../extdata/sample-rdf/fuseki-crnd-example-config.ttl` (see instruction in top of the file).

Alternative a virtouso endpoint can be used:
```{r, results='asis', eval=FALSE}
endpoint2<- "http://localhost:8890/sparql"
results.virtouso <- construct.remote(endpoint2, SPARQLscript )
summarize.rdf(results.virtouso)
```
This worked partially with virtuoso 7.2 as of 28-oct-2015: 164976 triples was retrieved. The fuseki endpoint was started using the configuration file `../extdata/sample-rdf/fuseki-crnd-example-config.ttl` (see instruction in top of the file). Note: the virtuoso endpoint must support remote retrival of data (see [http://localhost:8890/sparql?help=enable_sponge]) and the value of 
`ResultSetMaxRows` in the SPARQL section of `virtuoso.ini` should be sufficiently high; for the rdf.cdisc.org mentioned here a value of 200000 may be needed.

### Using the script with jena arq
The generated SPARQL query works with arq when using a local endpoint.
Replace _get-rdf-disc-org.rq_ with the path to the generated file (see above).

Execute the commands from a temporary directory (_/tmp_ is used below).

The script below will create a turtle file _construct-rdf.disc.org.ttl_ in the 
```
export JENAROOT=/opt/apache-jena-3.0.0
export PATH=$PATH:$JENAROOT/bin
arq --query=get-rdf-disc-org.rq > construct-rdf-disc-org.ttl
```

The next statements loads the turtle files generated by the arq invocation above.
```{r, results='asis', eval=FALSE}
arqCDISCTtl<- file.path("/tmp", "construct-rdf-disc-org.ttl")
cdisc.rdf.arq.0<- new.rdf(ontology=FALSE)
cdisc.rdf.arq<- load.rdf( arqCDISCTtl, format="TURTLE", appendTo=cdisc.rdf.arq.0)
summarize.rdf(cdisc.rdf.arq)
```
This worked as expected as of 08-mar-2016. 

However, the sparql query does not work exectured from the Apache Jena Fuseki Version 2.0 or Version 2.3.1. This may be due to some configuration option.


