---
title: How to create vignette"
author: "PhuseSubTeamAnalysisResults@example.org"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to create vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---


### How to run from command prompt when developing

Start in the package root directory

Devtools should reply reply Loading rrdfqbcrnd0.

The knit command should generate .md file in the present work directory - the package root directory


```{r, eval=FALSE}
library(devtools)
devtools::load_all()

library(knitr)
knit("vignettes/cube-from-workbook.Rmd")
```

Generate the pdf file using the following command
```{r makepdf.sh, eval=FALSE, engine='bash'}
pandoc qube-from-workbook.md -o qube-from-workbook.pdf
```

The knitr::knit2html should generate a .html file in the present work directory - the package root directory

Generating HTML works nicely
```{r, eval=FALSE} 
knitr::knit2html("vignettes/qube-from-workbook.Rmd")
```

The following does not work as the texi2dvi can not handle the dollar signs in the bash code above.
I suppose some other options have to be given. The work around is to use pandoc on the .md file generated as mentioned above.

```{r, eval=FALSE}
knitr::knit2pdf("vignettes/qube-from-workbook.Rmd")
```

