---
title: "Use R RDF version of CDISC standards"
author: "mja@statgroup.dk"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Create simple RDF data qube}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

## Introduction
This vignette shows how to 
* load the CDISC standards from the packages
* make a SPARQL query for extracting a codelist
* make a SPARQL query for extracting the definition for the SDTM DM dataset

## Setup

```{r, results='asis', eval=TRUE}
library(rrdfancillary)
library(rrdfcdisc)
```

### Loading using Load.cdisc.standards 

The code below shows the call for loading the zipped rdf file and store in a triple store.
This is currently handled in the .onLoad function for the package. The rrdf model is then stored in the package environment.

```{r, results='asis', eval=TRUE}
cdisc.rdf<- Load.cdisc.standards()
message(".. total number of triples: ", summarize.rdf.noprint(cdisc.rdf) )
```

### Extract codelist

```{r, results='asis', eval=TRUE}
nciDomainValue<- "sdtmct:C66731"
query <- GetCDISCCodeListSparqlQuery( Get.rq.prefixlist.df(qbCDISCprefixes), nciDomainValue )
codeSource <- as.data.frame(sparql.rdf(cdisc.rdf, query),stringsAsFactors=FALSE)

knitr::kable(codeSource)
```

### Extract the definition for the SDTM DM dataset

```{r, results='asis', eval=TRUE}
query.rq<-'
select *
where {
  ?column a <http://rdf.cdisc.org/mms#Column> .
  ?column <http://rdf.cdisc.org/mms#context> ?dataset .
  ?dataset a <http://rdf.cdisc.org/mms#Dataset>.
  optional { ?column <http://rdf.cdisc.org/mms#dataElementLabel> ?label }
  optional { ?column <http://rdf.cdisc.org/std/schema#controlledTermsOrFormat> ?controlledTermsOrFormat }
  optional { ?column <http://rdf.cdisc.org/std/schema#dataElementCompliance> ?dataElementCompliance }
  optional { ?column <http://rdf.cdisc.org/std/schema#dataElementRole> ?dataElementRole }
  optional { ?column <http://rdf.cdisc.org/std/schema#dataElementType> ?dataElementType }
  optional { ?column <http://rdf.cdisc.org/std/schema#references> ?references }
  optional { ?column <http://rdf.cdisc.org/mms#dataElement> ?dataElement }
  optional { ?column <http://rdf.cdisc.org/mms#dataElementDescription> ?dataElementDescription }
  optional { ?column <http://rdf.cdisc.org/mms#dataElementName> ?dataElementName }
  optional { ?column <http://rdf.cdisc.org/mms#dataElementValueDomain> ?dataElementValueDomain }
  optional { ?column <http://rdf.cdisc.org/mms#ordinal> ?ordinal }
  values (?dataset) {(<http://rdf.cdisc.org/std/sdtmig-3-1-3#Dataset.DM>)}
  }
  order by ?dataset ?ordinal 
'
datasetDef <- as.data.frame(sparql.rdf(cdisc.rdf, query.rq),stringsAsFactors=FALSE)
```

Here are some of the columns in the result:
```{r, results='asis', eval=TRUE}
knitr::kable(datasetDef[,c("dataset", "ordinal", "dataElementName", "label")])
```
