---
title: "Plot the indivdual subgraphs for use in the Tech Spec documentation"
author: "Tim.Williams@ucb.com"
date: '`r Sys.Date()`'
output: rmarkdown::html_vignette
---
 

# Program header


**DESCR** Plot the indivdual subgraphs for use in the Tech Spec documentation 
--Full Graph-- can be used as the graph in the Appendix. 

**NOTES** Unlike the pure D3 version, the RShiny App  does not  position the nodes 
and Unique node names are used instead of generating unique numeric  
ID values to connect nodes. 
Code is sloppy, but documented. Has warnings. No warranty implied. 

**INPUT** CubeStructureMetaData.xlsx , in same folder as the App. 

** ToDO** Fix warnings: 

1. "attributes are not identical across measure variables; they will be dropped"


```{r,eval=TRUE}

# install.packages("webshot")
library(webshot)
library(plyr) 
library(reshape2)
library(visNetwork)
library(xlsx)
library(shiny)
devtools::load_all(pkg="../.." )

```

```{r,eval=TRUE}
CubeStructureMetaDataFn<- system.file("extdata/spec-xls","CubeStructureMetaData.xlsx", package="rrdfqbspec")
cat("Reading specifications from ", normalizePath(CubeStructureMetaDataFn))

#------------------------------------------------------------------------------
# Nodes
#------------------------------------------------------------------------------
nodes <- read.xlsx(CubeStructureMetaDataFn,sheetName="Nodes") 

# Determine if duplicate node identifiers
#  Run this commented-out code whenever the spreadsheet is updated.
    # index <- duplicated(nodeList$node)
    # nodeList[index,]

# Remove duplicates before assigning unique ID number
nodes <- nodes[!duplicated(nodes$node),] 

# Recode the colors from colour group ID to hex color values.
nodes$color[nodes$nodeColor == "1"] <- "#FFFF00" # Yellow
nodes$color[nodes$nodeColor == "2"] <- "#989898" # Grey
nodes$color[nodes$nodeColor == "3"] <- "#FFFFFF" # White
nodes$color[nodes$nodeColor == "4"] <- "#377eb8" # Blue
nodes$color[nodes$nodeColor == "5"] <- "#006600" # Green
nodes$color[nodes$nodeColor == "6"] <- "#e41a1c" # Red
nodes$color[nodes$nodeColor == "7"] <- "#ff9933" # Orange  

# dot  = regular node
# star = links to other subgraphs in the display
# box  = center node of a subgraph. Is a selection in the drop down. Used to 
#        indicate subgraphs in the --Full Model-- graph 
nodes$shape[nodes$toOtherSubgraph == 0] <- "dot" # nodes$shape <- "dot"
nodes$shape[nodes$toOtherSubgraph == 1] <- "star" # 
nodes$shape[nodes$toOtherSubgraph == 2] <- "box" # Not used. Obscures parts of graph.

nodes$id <-nodes$node  # Rename as needed for the graph  
nodes$size <- 5 # Node sizes = all the same
nodes$label <- nodes$nodeLabel  # Or change this to a rename

#------------------------------------------------------------------------------
# Edges
#------------------------------------------------------------------------------
edges <- read.xlsx(CubeStructureMetaDataFn,sheetName="Edges") 


# Remove cases where graphName is NA
# Keep on only the cases where roleN is not NA. Consider use of another variable
naIndex <- complete.cases(edges[,"graphName"]) 
edges <- edges[naIndex, ]

# Rename for visNetwork
edges <- rename(edges, c("s" = "from", "o" = "to"))

# Set the line type for edges by recoding 'necessity' value
# If necessity is Optional, dashes=TRUe, else dashes=FALSE. Change to ifelse?
edges$dashes = FALSE
edges$dashes[edges$necessity == 'Optional'] <- TRUE

# Label for edges 
edges$label <- edges$p
# Add arrows to the TO destination node for each edge.
# Other graphs will be  more complex.
edges$arrows <- paste("to")
edges$width <- 2
# Edge label values
edges$align <- paste("top")

# List of subgraph names to use in the drop down selector
subGraphs <- sort(unique(as.character(edges$graphName))) 
subGraphs <- c(subGraphs, "--Full Model--")  # Add choice to include all nodes and edges


onegraph<- function(subGraphSelector) { 
    # Subset the edges dataset based on the graphName selected
    if(subGraphSelector == "--Full Model--"){
                edgesSub <- edges
            }
            else
            {
                edgesSub <- subset(edges, graphName == subGraphSelector)    
            }

    if(subGraphSelector =="--Full Model--"){
        edgesSub <- edges
    }
    else
    {
        edgesSub <- subset(edges, graphName == subGraphSelector)
    }
    
    edgeNodes <- edgesSub[,c("graphName", "from","to")]  ## Keep graphName as an IDVar
    ## Use the edgesSub to select only those nodes that are in edgesSub
    ## Get the lists of all the nodes in edgesSub (both source and target nodes)
    edgeNodes <- melt(edgeNodes, id.vars = "graphName")

    ## Remove duplicate node values (from repeats of the source node)
    edgeNodes <- edgeNodes[!duplicated(edgeNodes$value),]
    ## edgeNodes$value is the list of node names in edgesSub
    selRows <- nodes$node %in% edgeNodes$value  ## index of the matching rows
    nodesSub <- nodes[selRows,] ## subset to only those that match

    ## Arrows only on the "to" node
        visNetwork(nodesSub, edgesSub, 
               width = "100%",
               height = "600")  %>%
        visEdges(font = list(size = 16, face = 'sans-serif', color = "darkblue", align = "top"),
                 arrows = list(to = list(enabled = TRUE, scaleFactor = 0.5)),
                 color = "#ccc",
                 length = 250
                 ) %>%
        visNodes( font = list(size = 16, face = 'sans-serif', color = "black", background = "white"),
                 fixed = FALSE                  
                 )

}


```

```{r,eval=TRUE}
onegraph(subGraphs[1])
```

```{r,eval=TRUE}
onegraph(subGraphs[2])
```

```{r,eval=TRUE}
onegraph(subGraphs[3])
```

```{r,eval=TRUE}
onegraph(subGraphs[4])
```

```{r,eval=TRUE}
onegraph(subGraphs[5])
```

```{r,eval=TRUE}
onegraph(subGraphs[6])
```

```{r,eval=TRUE}
onegraph(subGraphs[7])
```

```{r,eval=TRUE}
onegraph(subGraphs[8])
```

```{r,eval=TRUE}
onegraph(subGraphs[9])
```

```{r,eval=TRUE}
onegraph(subGraphs[10])
```
