---
title: "Explore RDF data cube vocabulary"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Explore RDF data cube vocabulary}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
output:
  html_document:
    toc: true
    theme: united
  pdf_document:
    toc: true
    highlight: zenburn
  md_document:
    variant: markdown_github
---


# Explore RDF data cube vocabulary 

This script explores the RDF data cube vocabulary.

## Setup 
```{r}
library(rrdfancillary)
library(rrdfqb)
library("DiagrammeR")
```

The following is just a test for Mermaid.
Mermaid can also include links - see (http://knsv.github.io/mermaid/#interaction)
```{r, eval=FALSE}
mermaid.commands<- '
graph TB
 qb:Attachable("Attachable(abstract)")
 qb:Observation(Observation)
 qb:Attachable-->|text|qb:Observation
'
DiagrammeR( mermaid.commands )
```

# Load the vocabulary

The vocabulary is loaded in to a `rrdf` store with support for inferencing (`ontology=TRUE`).
Enabeling the inferencing is an advantage when loading an ontology/vocabulary like the RRDF data cube, as the vocabulary may be made expecting inferencing being used.

```{r, results='asis', eval=TRUE}
qb.fn <- file.path(system.file("extdata/cube-vocabulary-rdf", package="rrdfqb"), "cube.ttl" )
empty.rdf <- new.rdf(ontology=TRUE)
cat("Loading from ", normalizePath(qb.fn), "\n" )
qb.rdf<- load.rdf(qb.fn,"TURTLE",empty.rdf)
dump.qb.rq<- paste(
  "SELECT  ?s ?p ?o ",
  "WHERE { ?s ?p ?o }", sep="\n", collapse="\n"
)
cat(dump.qb.rq,"\n")
cube.triples<- data.frame(sparql.rdf(qb.rdf, dump.qb.rq ), stringsAsFactors=FALSE)
knitr::kable(cube.triples)
```

### Classes

First the data are loaded into a `rrdf` store. The code below uses `ontology=FALSE` as only the directly stated triples are of interest for this. 

```{r, results='asis', eval=TRUE}
empty.rdf <- new.rdf(ontology=FALSE)
qb.rdf<- load.rdf(qb.fn,"TURTLE",empty.rdf)
qb.prefix<- '
PREFIX rdf:            <http://www.w3.org/1999/02/22-rdf-syntax-ns#> 
PREFIX rdfs:           <http://www.w3.org/2000/01/rdf-schema#> 
PREFIX owl:            <http://www.w3.org/2002/07/owl#> 
PREFIX xsd:            <http://www.w3.org/2001/XMLSchema#> 
PREFIX skos:           <http://www.w3.org/2004/02/skos/core#> 
PREFIX foaf:           <http://xmlns.com/foaf/0.1/> 
PREFIX scovo:          <http://purl.org/NET/scovo#> 
PREFIX void:           <http://rdfs.org/ns/void#> 
PREFIX qb:             <http://purl.org/linked-data/cube#> 
PREFIX dcterms:        <http://purl.org/dc/terms/> 
'
```

Getting an overview of the classes is done as shown below, as suggested in "Learning SPARQL".

```{r, results='asis', eval=TRUE}
dump.classes.all.rq<- '
SELECT * 
WHERE 
{ 
  ?class a rdfs:Class .
  ?class a ?classType
}
'
cat(dump.classes.all.rq,"\n")
classes.all.df<- data.frame(sparql.rdf(qb.rdf, paste( qb.prefix, dump.classes.all.rq, sep="\n" )), stringsAsFactors=FALSE)
knitr::kable(classes.all.df)
```

For the present purpose only the class names are of interest.
```{r, results='asis', eval=TRUE}
dump.classes.distinct.rq<- '
SELECT DISTINCT ?class ?label
WHERE 
{ 
  ?class a rdfs:Class .
  ?class a ?classType .
  OPTIONAL {?class rdfs:label ?label}
}
'
cat(dump.classes.distinct.rq,"\n")
classes.distinct.df<- data.frame(sparql.rdf(qb.rdf, paste( qb.prefix, dump.classes.distinct.rq, sep="\n" )), stringsAsFactors=FALSE)
knitr::kable(classes.distinct.df)
```

For the present purpose only the class names are of interest.
```{r, results='asis', eval=TRUE}
dump.classes.subClassOf.rq<- '
SELECT DISTINCT ?class ?subClassOf 
WHERE 
{ 
  ?class a rdfs:Class .
  ?class rdfs:subClassOf ?subClassOf .
FILTER ( ?subClassOf != rdf:Property )
}
'
cat(dump.classes.subClassOf.rq,"\n")
classes.subClassOf.df<- data.frame(sparql.rdf(qb.rdf, paste( qb.prefix, dump.classes.subClassOf.rq, sep="\n" )), stringsAsFactors=FALSE)
knitr::kable(classes.subClassOf.df)
```

Finally, generate the `mermaid` commands for showing the class structure.
```{r, eval=TRUE}
mermaid.commands<- paste(
           "graph TB", 
           paste( classes.distinct.df$class,'("', classes.distinct.df$label, '")', sep="", collapse=";\n" ),
          paste( classes.subClassOf.df$subClassOf, "-->", classes.subClassOf.df$class, sep="", collapse=";\n" ),
          sep=" \n", collapse="\n")
cat(mermaid.commands,"\n")
```

Then display the graph.
```{r, eval=TRUE}
DiagrammeR( mermaid.commands )

```

## Domain and range with properties

The query below removes the blank nodes, as it only occurs for the `qb:codeList`.
```{r, results='asis', eval=TRUE}
dump.drp.rq<- '
SELECT DISTINCT ?domain ?property ?range # ( !isBlank(?range) as ?b)
WHERE 
{ 
  ?property a rdf:Property .
  ?property rdfs:domain ?domain .
  ?property rdfs:range ?range .
FILTER (  !isBlank(?range) )
}
'
cat(dump.drp.rq,"\n")
drp.df<- data.frame(sparql.rdf(qb.rdf, paste( qb.prefix, dump.drp.rq, sep="\n" )), stringsAsFactors=FALSE)
knitr::kable(drp.df)
```

Finally, generate the `mermaid` commands for showing the class structure.
```{r, eval=TRUE}
mermaid.commands<- paste(
           "graph TB", 
           paste( drp.df$domain,'("', drp.df$domain, '")', sep="", collapse=";\n" ),
           paste( drp.df$range,'("', drp.df$range, '")', sep="", collapse=";\n" ),
          paste( drp.df$domain, "-->", '|"', drp.df$property, '"|', drp.df$range, sep="", collapse=";\n" ),
          sep=" \n", collapse="\n")
cat(mermaid.commands,"\n")
```

Then display the graph.
```{r, eval=TRUE}
DiagrammeR( mermaid.commands )

```



```{r, eval=FALSE}
mermaid.commands<- '
graph TB 
qb:Slice("qb:Slice");
rdf:Property("rdf:Property");
qb:HierarchicalCodeList("qb:HierarchicalCodeList");
qb:HierarchicalCodeList-->|"qb:parentChildProperty"|rdf:Property;
qb:DataSet("qb:DataSet");
qb:Observation("qb:Observation");
qb:ComponentProperty("qb:ComponentProperty");
qb:ObservationGroup("qb:ObservationGroup");
qb:DataSet("qb:DataSet");
qb:ComponentSpecification("qb:ComponentSpecification");
qb:DataStructureDefinition("qb:DataStructureDefinition");
qb:ComponentSet("qb:ComponentSet") 
qb:SliceKey("qb:SliceKey");
rdfs:Class("rdfs:Class");
qb:DataStructureDefinition("qb:DataStructureDefinition");
xsd:int("xsd:int");
qb:DataSet("qb:DataSet");
skos:Concept("skos:Concept");
qb:Observation("qb:Observation");
qb:Slice("qb:Slice");
xsd:boolean("xsd:boolean");
qb:SliceKey("qb:SliceKey");
qb:ComponentProperty("qb:ComponentProperty") 
qb:Slice-->|"qb:sliceStructure"|qb:SliceKey;
qb:ComponentSpecification-->|"qb:componentAttachment"|rdfs:Class;
qb:DataSet-->|"qb:structure"|qb:DataStructureDefinition;
qb:ComponentSpecification-->|"qb:order"|xsd:int;
qb:Observation-->|"qb:dataSet"|qb:DataSet;
qb:ComponentProperty-->|"qb:concept"|skos:Concept;
qb:ObservationGroup-->|"qb:observation"|qb:Observation;
qb:DataSet-->|"qb:slice"|qb:Slice;
qb:ComponentSpecification-->|"qb:componentRequired"|xsd:boolean;
qb:DataStructureDefinition-->|"qb:component"|qb:ComponentSpecification;
qb:DataStructureDefinition-->|"qb:sliceKey"|qb:SliceKey;
qb:ComponentSet-->|"qb:componentProperty"|qb:ComponentProperty
'

DiagrammeR( mermaid.commands )

```
