---
title: "RDF data cube SPARQL queries"
vignette: >
  %\VignetteIndexEntry{Explore RDF data cube vocabulary}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
output:
  html_document:
    toc: true
    theme: united
  pdf_document:
    toc: true
    highlight: zenburn
  md_document:
    variant: markdown_github
---

# Setup
```{r, results='asis', eval=TRUE}
library(rrdfancillary)
library(rrdfqb)
```

# Load example cube and define SPARQL prefixes

```{r, results='asis', eval=TRUE}
qbfile <- file.path(system.file("extdata/sample-rdf", "example-normalized.ttl", package="rrdfqb") )
# qbfile <- file.path(system.file("extdata/sample-rdf", "example.ttl", package="rrdfqb") )
# qbfile <- file.path(system.file("extdata/sample-rdf", "DC-DEMO-sample.ttl", package="rrdfcrndex") )
cube <- new.rdf(ontology=FALSE)
invisible(load.rdf( qbfile, format="TURTLE", cube))

SPARQLprefix<- '
prefix rdf: <http://www.w3.org/1999/02/22-rdf-syntax-ns#>  
prefix skos: <http://www.w3.org/2004/02/skos/core#>  
prefix prov: <http://www.w3.org/ns/prov#>  
prefix rdfs: <http://www.w3.org/2000/01/rdf-schema#>  
prefix dcat: <http://www.w3.org/ns/dcat#>  
prefix owl: <http://www.w3.org/2002/07/owl#>  
prefix xsd: <http://www.w3.org/2001/XMLSchema#>  
prefix qb: <http://purl.org/linked-data/cube#>  
prefix pav: <http://purl.org/pav>  
prefix dct: <http://purl.org/dc/terms/>  
'
```

# Get data set

```{r, results='asis', eval=TRUE}

SPARQLquery<- '
select ?dataset
where {
?dataset a qb:DataSet .
} 
'

SPARQLscript<- paste( SPARQLprefix, SPARQLquery )
result <- sparql.rdf(cube, SPARQLscript )
knitr::kable(result)

```

# Get the data structure definition in the store

```{r, results='asis', eval=TRUE}

SPARQLquery<- '
select ?dataset ?dsd
where {
?dataset a qb:DataSet ;
  qb:structure ?dsd .
 ?dsd a qb:DataStructureDefinition  .
} 
'

SPARQLscript<- paste( SPARQLprefix, SPARQLquery )
result <- sparql.rdf(cube, SPARQLscript )
knitr::kable(result)

```

# Get the componentes in the data structure definition in the store

```{r, results='asis', eval=TRUE}

SPARQLquery<- '
select ?dataset ?dsd ?component ?componentPropery ?componentReference 
where {
?dataset a qb:DataSet ;
  qb:structure ?dsd .
?dsd a qb:DataStructureDefinition;
qb:component ?component .
?component ?componentPropery ?componentReference .
} 
'

SPARQLscript<- paste( SPARQLprefix, SPARQLquery )
result <- sparql.rdf(cube, SPARQLscript )
knitr::kable(result)

```

# Get the dimension, measures and attributes

```{r, results='asis', eval=TRUE}
SPARQLquery<- '
select ?dataset ?component ?isblankcomponent ?componentProperty ?componentReference ?ordercp ?order ?label
where {
?dataset a qb:DataSet ;
  qb:structure ?dsd .
?dsd a qb:DataStructureDefinition;
qb:component ?component .
?component ?componentProperty ?componentReference .
optional{ ?component qb:order ?temporder . }
BIND( ISBLANK(?component) as ?isblankcomponent )
optional{ ?componentReference rdfs:label ?templabel . }
BIND(
if( ?componentProperty = qb:dimension, 1001, 
   if( ?componentProperty = qb:measure, 2001,
      if( ?componentProperty = qb:attribute, 3001, 9001 )
        )
  )
 as ?ordercp )
BIND( COALESCE( ?temporder, 0 )  as ?order )
BIND( COALESCE( ?templabel, "" )  as ?label )
values ( ?componentProperty ) {
(qb:attribute)
(qb:dimension)
(qb:measure)
}
}
order by ?dataset ?ordercp ?order ?componentReference
'
SPARQLscript<- paste( SPARQLprefix, SPARQLquery )
result <- data.frame(sparql.rdf(cube, SPARQLscript ),stringsAsFactors=FALSE)
knitr::kable(result)
```

Add CONSTRUCT query that defines the cube from the SPARQL results above.

# Make SPARQL query for getting observations for example.ttl

This will not work as the cube is not normalized.


```{r, results='asis', eval=FALSE}
genfor<- result[result$dataset==result$dataset[1],]
genfor$colnames<- sapply(strsplit(genfor$componentReference, "#"), "[", 2)

SPARQLquery<- paste( "select * where {",
      "?obs a qb:Observation;",
      "# qb:dataSet  ?dataset ; ",
      paste("",
          "<", genfor$componentReference, ">", " ", "?", genfor$colnames, " ;", sep="", collapse="\n"
      ),
      ".", 
      paste( "#values (?dataset) { (",
            "<", result$dataset[1], ">",
            ") }", sep="", collapse="\n"
            ),
      "}", sep="\n", collapse="\n"
      )

SPARQLscript<- paste( SPARQLprefix, SPARQLquery )
cat(SPARQLscript, "\n")
oresult <- data.frame(sparql.rdf(cube, SPARQLscript ),stringsAsFactors=FALSE)
knitr::kable(oresult)



```

# Make SPARQL query for getting observations for example.ttl

This query below expects the cube is normalized, which is the case for
the cube loaded in the store. Note, it will not work on the example
cube as presented in RDF Data Cube Vocabulary, as the example cube is
by purpose not normalised.


```{r, results='asis', eval=TRUE}
genfor<- result[result$dataset==result$dataset[1],]
genfor$colnames<- sapply(strsplit(genfor$componentReference, "#"), "[", 2)

SPARQLquery<- paste( "select * where {",
      "?obs a qb:Observation;",
      "qb:dataSet  ?dataset ; ",
      paste(
          "<", genfor$componentReference, ">", " ", "?", genfor$colnames, " ;", sep="", collapse="\n"
      ),
      ".", 
      paste( "values (?dataset) { (",
            "<", result$dataset[1], ">",
            ") }", sep="", collapse="\n"
            ),
      "}", sep="\n", collapse="\n"
      )

SPARQLscript<- paste( SPARQLprefix, SPARQLquery )
cat(SPARQLscript, "\n")
oresult <- data.frame(sparql.rdf(cube, SPARQLscript ),stringsAsFactors=FALSE)
knitr::kable(oresult)
```
