---
title: "SQL code for verifying results in RDF data cube"
author: "mja@statgroup.dk"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{SQL code for verifying results in RDF data cube}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

# Derive results in RDF data cube and compare with results in data cube

## Setup
Here all libraries are loaded; this should not be necessary.
```{r}
options(width=200) # long lines
library(xlsx)
library(foreign)
library(rrdf)
library(rrdfqb)
library(rrdfqbcrnd0)
library(rrdfqbcrndex)
library(rrdfqbcrndcheck)
```

## Load the data
The SAS dataset is loaded, and factors converted to character.
```{r}
obsFile<- system.file("extdata/sample-xpt", "adsl.xpt", package="rrdfqbcrndex")
adsl<-read.xport(obsFile)
ii <- sapply(adsl, is.factor)
adsl[ii] <- lapply(adsl[ii], as.character)
```

The conversion to character can be avoided by using
`library(SASxport)`, see
(../../rrdfqbcrndex/inst/data-raw/create-dm-table-as-csv.Rmd).

## Load the RDF data cube
The RDF data cube is loaded.
```{r}
dataCubeFile<- system.file("extdata/sample-rdf", "DC-DEMO-sample.ttl", package="rrdfqbcrndex")
store <- new.rdf()  # Initialize
cat("Reading turtle definition from ", dataCubeFile, "\n")
temp<- load.rdf(dataCubeFile, format="TURTLE", appendTo= store)
summarize.rdf(store)
```

With a SPARQL query the mean values for the `WEIGHTBL` variable in RDF data cube the are extracted.
```{r}
rq<-'
prefix crnd-measure: <http://www.example.org/dc/measure#> 
prefix code:  <http://www.example.org/dc/code/> 
prefix crnd-attribute: <http://www.example.org/dc/attribute#> 
prefix ds:    <http://www.example.org/dc/demo/ds/> 
prefix qb:    <http://purl.org/linked-data/cube#> 
prefix crnd-dimension: <http://www.example.org/dc/dimension#> 
select * where {
?obs  a                         qb:Observation ;
    qb:dataSet                  ds:dataset-DEMO ;
        crnd-dimension:agegr1       code:agegr1-_ALL_ ;
        crnd-dimension:ethnic       code:ethnic-_ALL_ ;
        crnd-dimension:factor       code:factor-weightbl ;
        crnd-dimension:procedure    code:procedure-mean ;
        crnd-dimension:race         code:race-_ALL_ ;
        crnd-dimension:sex          code:sex-_ALL_ ;
        crnd-dimension:trt01a       ?trt01a ;
    crnd-measure:measure        ?measure .
}
'
knitr::kable(data.frame(sparql.rdf( store, rq)))
```

The SPARQL query about can be made more generic. 

## Generate SQL statements
From the RDF data cube loading in the store, the function `GetSQLFromCube` generates the SQL statements for reproducing the data cube.
Only the first two select statements are shown.

```{r}
stmtSQL<- GetSQLFromCube( store, srcDsName="adsl" )
cat(paste(unlist(strsplit(stmtSQL$summStatSQL,split="\n"))[1:3],collapse="\n"),"\n")
```

## Derive the descriptive statistics

To show the full SQL expression use

```{r,eval=FALSE}
cat(stmtSQL$summStatSQL)
```

```{r}
library(sqldf)
adsl.summ.stat.res<- sqldf( stmtSQL$summStatSQL)
names(adsl.summ.stat.res)<- tolower(gsub("(a|b)\\.","", names(adsl.summ.stat.res)))
knitr::kable(adsl.summ.stat.res)
```

## Quick check
Sometimes it is usefull to see that there is an actual derivation going on.
The calculated mean values for the `WEIGHTBL` variable is
```{r}
knitr::kable(adsl.summ.stat.res[adsl.summ.stat.res$factor=="WEIGHTBL" & adsl.summ.stat.res$procedure=="mean",])
```

If `WEIGHTBL` is multiplied by 100, then mean should also be multiplied with 100:
```{r}
adsl$WEIGHTBL<- adsl$WEIGHTBL*100
adsl.summ.stat.mod.res<- sqldf( stmtSQL$summStatSQL)
knitr::kable(adsl.summ.stat.mod.res[adsl.summ.stat.mod.res$factor=="WEIGHTBL" & adsl.summ.stat.mod.res$procedure=="mean",])
```
